### 2018-10-07 paper/e511-proposal.tex
Project proposal documents.

% This is "sig-alternate.tex" V2.0 May 2012
% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
%
% This example file demonstrates the use of the 'sig-alternate.cls'
% V2.5 LaTeX2e document class file. It is for those submitting
% articles to ACM Conference Proceedings WHO DO NOT WISH TO
% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
% The 'sig-alternate.cls' file will produce a similar-looking,
% albeit, 'tighter' paper resulting in, invariably, fewer pages.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V2.5) produces:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) NO page numbers
%
% as against the acm_proc_article-sp.cls file which
% DOES NOT produce 1) thru' 3) above.
%
% Using 'sig-alternate.cls' you have control, however, from within
% the source .tex file, over both the CopyrightYear
% (defaulted to 200X) and the ACM Copyright Data
% (defaulted to X-XXXXX-XX-X/XX/XX).
% e.g.
% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
%
% ---------------------------------------------------------------------------------------------------------------
% This .tex source is an example which *does* use
% the .bib file (from which the .bbl file % is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% ================= IF YOU HAVE QUESTIONS =======================
% Questions regarding the SIGS styles, SIGS policies and
% procedures, Conferences etc. should be sent to
% Adrienne Griscti (griscti@acm.org)
%
% Technical questions _only_ to
% Gerald Murray (murray@hq.acm.org)
% ===============================================================
%
% For tracking purposes - this is V2.0 - May 2012

\documentclass{sig-alternate}

\begin{document}
%
% --- Author Metadata here ---
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{Project Proposal: Monaural Single-Source Speaker Separation}
\subtitle{E511 Machine Language for Signal Processing}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{3} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Jonathan Branam
       \email{jobranam@iu.edu}
% 2nd. author
\alignauthor
Jamie Israel
       \email{jgisrael@iu.edu}
% 3rd. author
\alignauthor
Hellen Jiang
       \email{heljiang@iu.edu}
}

\maketitle
\begin{abstract}
There have been several recent developments in the use of deep
learning to perform speaker diarisation, or separation of multiple
speech signals from a single source into separate homogenous signals
representing each individual speaker.  This task is particularly
challenging where the input is monophonic, there is no control over
environmental conditions, the number of speakers is unknown and there
are no samples of the speakers from which to train. One such
real-world application is for a call center seeking to analyze the
speech of agents and customers where the only source is a monophonic
recording of a call with unknown participants.  The center would like
to transcribe speech to text for content analysis, but they have been
unable to obtain an accurate translation due in large part to overlap
in the recorded signals (i.e., cross-talk) and the inability to
associate utterances with speakers.

Our initial step will be to apply the non-negative matrix
factorization technique illustrated in the class materials to perform
a baseline for this source separation problem. From there, we intend
to explore a neural network model using PyTorch to implement
Permutation Invariant Training (PIT) from
\cite{DBLP:journals/corr/KolbaekYTJ17} or utterance-level Permutation
Invariant Training (uPIT) from
\cite{DBLP:journals/corr/YuKTJ16}
as these seem to be the latest approaches to
single source sound separation in the academic literature. The call
center recordings are not currently labeled, so we propose creating
our own test and training dataset by mixing single-speaker recordings
from either the TIMIT or WSJ0 recordings. The literature mentions the
WSJ0-2mix data set that was used in
\cite{DBLP:journals/corr/HersheyCRW15} the seminal Deep Clustering
(DPCL) paper but we have yet to find a specification for how to
generate this dataset.

\end{abstract}

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{e511_proposal}
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
\end{document}

### 2018-10-07 paper/e511-proposal.tex
Project proposal.

@@
 
-\documentclass{sig-alternate}
+\documentclass{e511-project}
 
@@
 as these seem to be the latest approaches to
-single source sound separation in the academic literature. The call
+single source speaker separation in the academic literature. The call
 center recordings are not currently labeled, so we propose creating
 our own test and training dataset by mixing single-speaker recordings
-from either the TIMIT or WSJ0 recordings. The literature mentions the
-WSJ0-2mix data set that was used in
+from either the TIMIT \cite{garofolo:timit} or WSJ0 
+\cite{garofolo:wsj0} recordings.
+The literature mentions the WSJ0-2mix data set that was used in
 \cite{DBLP:journals/corr/HersheyCRW15} the seminal Deep Clustering
-(DPCL) paper but we have yet to find a specification for how to
+(DPCL) paper but we have yet to find a detailed specification for how to
 generate this dataset.
@@
 \bibliographystyle{abbrv}
-\bibliography{e511_proposal}
+\bibliography{e511-proposal}
 % You must have a proper ".bib" file

### 2018-12-13 paper/main.tex
Backup paper.

\documentclass[journal, a4paper]{IEEEtran}

% some very useful LaTeX packages include:

%\usepackage{cite}      % Written by Donald Arseneau
% * <yueguo@iu.edu> 2018-09-01T13:39:33.474Z:
%
% ^.
                        % V1.6 and later of IEEEtran pre-defines the format
                        % of the cite.sty package \cite{} output to follow
                        % that of IEEE. Loading the cite package will
                        % result in citation numbers being automatically
                        % sorted and properly "ranged". i.e.,
                        % [1], [9], [2], [7], [5], [6]
                        % (without using cite.sty)
                        % will become:
                        % [1], [2], [5]--[7], [9] (using cite.sty)
                        % cite.sty's \cite will automatically add leading
                        % space, if needed. Use cite.sty's noadjust option
                        % (cite.sty V3.8 and later) if you want to turn this
                        % off. cite.sty is already installed on most LaTeX
                        % systems. The latest version can be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/contrib/supported/cite/

\usepackage{graphicx}   % Written by David Carlisle and Sebastian Rahtz
                        % Required if you want graphics, photos, etc.
                        % graphicx.sty is already installed on most LaTeX
                        % systems. The latest version and documentation can
                        % be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/required/graphics/
                        % Another good source of documentation is "Using
                        % Imported Graphics in LaTeX2e" by Keith Reckdahl
                        % which can be found as esplatex.ps and epslatex.pdf
                        % at: http://www.ctan.org/tex-archive/info/

%\usepackage{psfrag}    % Written by Craig Barratt, Michael C. Grant,
                        % and David Carlisle
                        % This package allows you to substitute LaTeX
                        % commands for text in imported EPS graphic files.
                        % In this way, LaTeX symbols can be placed into
                        % graphics that have been generated by other
                        % applications. You must use latex->dvips->ps2pdf
                        % workflow (not direct pdf output from pdflatex) if
                        % you wish to use this capability because it works
                        % via some PostScript tricks. Alternatively, the
                        % graphics could be processed as separate files via
                        % psfrag and dvips, then converted to PDF for
                        % inclusion in the main file which uses pdflatex.
                        % Docs are in "The PSfrag System" by Michael C. Grant
                        % and David Carlisle. There is also some information
                        % about using psfrag in "Using Imported Graphics in
                        % LaTeX2e" by Keith Reckdahl which documents the
                        % graphicx package (see above). The psfrag package
                        % and documentation can be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/contrib/supported/psfrag/

%\usepackage{subfigure} % Written by Steven Douglas Cochran
                        % This package makes it easy to put subfigures
                        % in your figures. i.e., "figure 1a and 1b"
                        % Docs are in "Using Imported Graphics in LaTeX2e"
                        % by Keith Reckdahl which also documents the graphicx
                        % package (see above). subfigure.sty is already
                        % installed on most LaTeX systems. The latest version
                        % and documentation can be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/contrib/supported/subfigure/

\usepackage{url}        % Written by Donald Arseneau
                        % Provides better support for handling and breaking
                        % URLs. url.sty is already installed on most LaTeX
                        % systems. The latest version can be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/contrib/other/misc/
                        % Read the url.sty source comments for usage information.

%\usepackage{stfloats}  % Written by Sigitas Tolusis
                        % Gives LaTeX2e the ability to do double column
                        % floats at the bottom of the page as well as the top.
                        % (e.g., "\begin{figure*}[!b]" is not normally
                        % possible in LaTeX2e). This is an invasive package
                        % which rewrites many portions of the LaTeX2e output
                        % routines. It may not work with other packages that
                        % modify the LaTeX2e output routine and/or with other
                        % versions of LaTeX. The latest version and
                        % documentation can be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/contrib/supported/sttools/
                        % Documentation is contained in the stfloats.sty
                        % comments as well as in the presfull.pdf file.
                        % Do not use the stfloats baselinefloat ability as
                        % IEEE does not allow \baselineskip to stretch.
                        % Authors submitting work to the IEEE should note
                        % that IEEE rarely uses double column equations and
                        % that authors should try to avoid such use.
                        % Do not be tempted to use the cuted.sty or
                        % midfloat.sty package (by the same author) as IEEE
                        % does not format its papers in such ways.

\usepackage{amsmath}    % From the American Mathematical Society
                        % A popular package that provides many helpful commands
                        % for dealing with mathematics. Note that the AMSmath
                        % package sets \interdisplaylinepenalty to 10000 thus
                        % preventing page breaks from occurring within multiline
                        % equations. Use:
%\interdisplaylinepenalty=2500
                        % after loading amsmath to restore such page breaks
                        % as IEEEtran.cls normally does. amsmath.sty is already
                        % installed on most LaTeX systems. The latest version
                        % and documentation can be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% Other popular packages for formatting tables and equations include:

%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty which improves the
% LaTeX2e array and tabular environments to provide better appearances and
% additional user controls. array.sty is already installed on most systems.
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

% V1.6 of IEEEtran contains the IEEEeqnarray family of commands that can
% be used to generate multiline equations as well as matrices, tables, etc.

% Also of notable interest:
% Scott Pakin's eqparbox package for creating (automatically sized) equal
% width boxes. Available 
% http://www.ctan.org/tex-archive/macros/latex/contrib/supported/eqparbox/

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
\usepackage{listings}

% Your document starts here!
\begin{document}

% Define document title and author
	\title{E511-Signal Processing-Final Project}
	\author{Jonathan Branam, Jamie Israel, Jinju(Hellen) Jiang}
	\date{December, 13 2018}
	\maketitle
%
% ^.
% Write abstract here
\begin{abstract}
In this project we evaluated and implemented state of the art techniques to solve the monaural speech separation (diarization) problem using a variety of approached. We investigated classification, masking, and deep learning approaches to separate a generated set of mixed speech recordings.
\end{abstract}

% Each section begins with a \section{title} command
\section{Introduction}
There have been several recent developments in the use of deep learning to perform speaker diarization, or separation of multiple speech signals from a single source into separate homogeneous signals representing each individual speaker. This task is particularly challenging where the input is monophonic, there is no control over environmental conditions, the number of speakers is unknown and there are no samples of the speakers from which to train. Using the CSR-I(WSJ0) Complete dataset\cite{WSJintro}, we tested several algorithms for accomplishing monaural speech separation including classification (KMeans), various masking approaches (NMF, Ideal Ratio Mask[IRM], and complex Ideal Ratio Mask [cIRM]), and permutation-based deep learning approaches (Permutation Invariant Training [PIT] and utterance-level Permutation Invariant Training [uPIT]).

The motivating application is to separate agent and customer utterances from call center recordings, which involve short duration speech with occasional cross talk and at least one unknown speaker.

\section{Literature Review}

\subsection{Jamie's Lit Review}

\subsection{Deep Learning Approaches}

A seminal paper in applying deep learning to the task of speaker diarization was the Deep Clustering \cite{DBLP:journals/corr/HersheyCRW15} approach proposed by Hershey, et. al. They divided the task of separating source signals into two steps. The first step is to use a deep neural network to learn a set of embeddings that produce a class-independent, low-rank approximation of the sources. These embeddings are trained to minimize the distance between embeddings in the same partition while maximizing the distances between embeddings in different partitions. The partitions used in their technique do not include class labels so the model does not assign a particular class to each embedding. The resulting embeddings can then be separated using simple clustering algorithms such as $k$-means. This is in contrast to previous work that relies on \textit{spectral clustering} for segmentation which uses local affinity measures to optimize an objective function using spectral decomposition. The approach of Hershey, et. al. is typical in current research for deep learning: rather than using a complicated set of specially designed features the deep clustering approach uses a deep neural network to discover the best features for producing the desired partitions.

Results:

\section{Dataset Preparation}
\subsection{CSR-I(WSJ0) Complete}
For purposes of comparison with existing scholarship, each approach was developed based on the WSJ0 dataset, a corpus of Wall Street Journal text data organized by speakers.

\subsection{WSJ0-mix}

Developed based on...\cite{WSJmix}.

\section{Speech Separation Algorithms}
\subsection{KMeans}
The technique of speaker diarization relies on a big pipeline with following steps:
\begin{itemize}
    \item Feature Extraction
    \item Speaker segmentation
    \item Speaker Clustering
    \item Evaluation
\end{itemize}
\subsubsection{Feature Extraction}
In this project, we used 3 approaches to extract features:
\begin{itemize}
    \item A chroma vector (Wikipedia) (FMP, p. 123) is a typically a 12-element feature vector indicating how much energy of each pitch class, {C, C\#, D, D\#, E, ..., B}, is present in the signal. Then use mean of each element as one of input for clustering.
    \item The mel frequency cepstral coefficients (MFCCs) of a signal are a small set of features  which concisely describe the overall shape of a spectral envelope. In this case, mfcc computed 13 MFCCs.The very first MFCC, the 0th coefficient, does not convey information relevant to the overall shape of the spectrum. It only conveys a constant offset, i.e. adding a constant value to the entire spectrum. Therefore, here I discard the first MFCC, then I scale the MFCCs such that each coefficient dimension has zero mean and unit variance.Then use mean of MFCCS as one part of input for clustering.
    \item We divide the audio signal into smaller frames. In smaller time scales audio signals are statistically unchanged. Then for each smaller frames we compute the power spectrum of the signal, and use it as the third part of input for clustering.
\end{itemize}
\subsubsection{Speaker segmentation}\cite{Segmentation} Speaker segmentation which is also known as acoustic change detection aims to detect speaker change such that each contiguous segment corresponds to single speaker only. To find if the two segments correspond to same speaker we have to define some notion of distance metric. In this case, I skipped this step and used the features extrated from above steps as input for kmeans clustering.
\subsubsection{Kmeans Clustering}
Since we knew that the audio from call center is the conversation between 2 speakers(customer service and customer), we set k=2 for kmeans clustering.\\
here it is the results for speaker diarization.
\begin{figure}[h!]
    \centering  
     \caption{\label{Fig:speaker diarization}2 speakers diarization}  
    \includegraphics[width=0.4\textwidth]{kmeans01.png} 
\end{figure}

\begin{figure}[h!]
    \centering  
     \caption{\label{Fig:speaker diarization}2 speakers diarization}  
    \includegraphics[width=0.4\textwidth]{kmeans02.png}
\end{figure}
\subsubsection{Evaluation\cite{Segmentation}}Diarization Error Rate (DER) is used for evaluation of automatic speech recognition system.Contribution to DER comes from three factors namely, Missed speech rate (MSR), False alarm speech rate (FASR) and Speaker Error. When a speech is labeled as non-speech then that error comes under MSR. FASR is when a non- speech is detected as a speech segment. Speaker error is contributed due to speaker clustering and segmentation. This kind of error can be caused if a speaker change is not detected, oversegmentation, erroneously clustered. Sum of all three errors contribute to the DER.\\
In this case, we don't have any labeled data to calculate the training error, we just simply used human ears to justify whether the diarization was good or not. For some training data, it works well, but some training data, it performed really bad.

\subsection{NMF, IRM and cIRM}
Nonnegative matrix factorization (NMF) is used to represent high-dminesional data as the product of two matrices (typically referenced as \textit{W} and \textit{H}). In the case of an audio signal, these matrices can be viewed as a representation of the signal's frequency spectrum (\textit{W}) and the corresponding activations (\textit{H}).

To perform NMF, we transformed isolated audio samples of two speakers (speaker 1 and speaker 2) to a frequency-time representation using short-time Fourier transform (STFT) before decomposing each signal using the following iterative update rules:
\newline

$W = W \bigodot \frac{\frac{X}{WH}H^T}{1{FxT}H^T}$ \hspace{1cm} $H = H \bigodot \frac{W^T\frac{X}{WH}}{W^T1^{FxT}}$
\newline

Using the frequency matrices associated with the sample from each speaker, a new set of activations (\textit{W}) was generated from the mixed signal composed of both individual samples. A magnitude masking matrix, reflecting the magnitudes associated with the frequency representation for each speaker at each time frame, was generated using this new set of activations and the following formula:

\begin{center}
$W_{S1}H_{(1:30,:)}$

\rule[1pt]{100pt}{.4pt}

$W_{S1}H_{(1:30,:)}+W_{S2}H_{(31:60,:)}  $

\end{center}

\begin{flushleft}
where S1 is the basis vector associated with the speaker 1 audio sample, S2 is the basis vector associated with the speaker 2 audio sample and H is the activation matrix generated from the combined speaker 1 and speaker 2 basis vectors.\cite{ClassNMF}
\end{flushleft}

Using the STFT of a single speaker audio sample (speaker 1) and the STFT of the mixed audio sample that included speaker 1 [Figure 1], an IRM was generated based on the formula:
\begin{center}
$S_{(t,f)}^2$

\rule[1pt]{40pt}{.4pt}

$S_{(t,f)}^2 + N_{(t,f)}^2  $

\end{center}

\begin{flushleft}

where S is the STFT generated from the isolated speaker 1 audio signal and N is the STFT generated fromm the mixed audio signal that included the same audio sample of speaker 1.\cite{DBLP:journals/corr/abs-1708-07524}

\end{flushleft}

\begin{figure}[h!]
    \centering  
     \caption{\label{Fig:Spectrogram of isolated amd mixed audio sample from speaker 1}Spectrogram of isolated amd mixed audio sample from speaker 1}  
    \includegraphics[width=0.4\textwidth]{cIRM_same.png}  
\end{figure}

Similarly, a cIRM was generated from the audio sample of speaker 1 and the same mixed audio sample using the formula:

\begin{center}
$Y_r S_r + Y_i S_i \hspace{1.6cm} Y_r S_i - Y_i S_r$

\rule[3pt]{45pt}{.4pt} \hspace{.3cm} + \hspace{.4cm} \textit{i} \hspace{.1cm} \rule[3pt]{45pt}{.4pt}

$Y_r^2 + Y_i^2  \hspace{2.0cm} Y_r^2 + Y_i^2$

\end{center}

\begin{flushleft}
where S is the complex STFT representation generated from the isolated speaker 1 audio signal and Y is the complex STFT generated form the mixed audio signal that included the same audio sample of speaker 1.\cite{DBLP:journals/corr/abs-1708-07524}
\end{flushleft}

\subsection{PIT and uPIT}

..
.... Add text here....

\section{Results}

\subsection{Masking Methods}

With the NMF method, we were unable to generate a substantial separation between the original isolated signal and the mixed audio. Both IMR approaches produced substantially better results with the cIMR method proving to be the best method for this type of separation on a consistent basis [Figure 2].

\begin{figure}[h!]
    \centering  
     \caption{\label{Fig:Comparison of Average SNR for Masking Methods 1}Comparison of Average SNR for Masking Methods}  
    \includegraphics[width=0.3\textwidth]{Mask_Res.png}  
\end{figure}

\section{Conclusion}
...
\newline 
% Now we need a bibliography:

\bibliographystyle{plain} % We choose the "plain" reference style
\bibliography{report} % Entries are in the "refs.bib" file

% Your document ends here!
\end{document}

### 2018-12-13 paper/main.tex
Lit review moving on to PIT.

@@
 
-A seminal paper in applying deep learning to the task of speaker diarization was the Deep Clustering \cite{DBLP:journals/corr/HersheyCRW15} approach proposed by Hershey, et.
+A seminal paper in applying deep learning to the task of speaker diarization was the Deep Clustering (DPCL) \cite{DBLP:journals/corr/HersheyCRW15} approach proposed by Hershey, et.
 al.
 They divided the task of separating source signals into two steps.
-The first step is to use a deep neural network to learn a set of embeddings that produce a class-independent, low-rank approximation of the sources.
+The first step is to use a deep neural network to learn a set of embeddings that estimate a class-independent, low-rank pairwise affinity matrix.
 These embeddings are trained to minimize the distance between embeddings in the same partition while maximizing the distances between embeddings in different partitions.
@@
 al.
-is typical in current research for deep learning: rather than using a complicated set of specially designed features the deep clustering approach uses a deep neural network to discover the best features for producing the desired partitions.
+is typical in current research for deep learning: rather than using a complicated set of specially designed features, the deep clustering approach uses a deep neural network to discover the best features for producing the desired partitions.
 
-Results:
+This work is also important because it introduced a new training dataset for the task of monaural speech separation.
+The authors found that existing datasets were too limited in size or were not well fitted to their task.
+The deep learning techniques also require a large training set and so the authors adapted the Wall Street Journal (WSJ0) \cite{WSJintro} single-channel speaker corpus to produce a new dataset called WSJ0-2mix \cite{WSJmix} that was used by the later literature for comparison with DPCL.
+
+The Deep Clustering technique achieved Signal-to-Distortion (SDR) improvements of up to 6 dB which outperformed their baseline of an oracle NMF approach of 5 dB SDR which was trained on the clean speaker sources.
+It is worth noting the these initial results were improved to 10.3 dB by further optimization of the model by using a deeper architecture, a longer temporal context, and improving the regularization of the network by \cite{DBLP:journals/corr/IsikRCWH16}.
+This advancement shows the improvements possible by tuning hyper-parameters of a deep learning model when applied to the task of speaker diarization.
+
+One of the biggest problems plaguing deep learning techniques in source separation is the \textit{label permutation problem} which is that most separation algorithms must provide the reference magnitudes to the output layers of the network to compute the error.
+The ordering of these source streams is ambiguous and imposes an outside and unnecessary constraint on the training process.
+Note that DPCL, discussed above, avoids this problem due to its treatment of the source separation problem as a partitioning problem.
+Although DPCL was successful, it made a simplifying assumption that each time-frequency bin was assigned to a single speaker (determined by the results of clustering).
+In \cite{DBLP:journals/corr/YuKTJ16} Yu, et.
+al.
+proposed a new technique named permutation invariant training (PIT) that avoids the label ambiguity problem by treating the reference source streams as a set rather than an ordered list.
 

### 2018-12-13 paper/main.tex
PIT summary.

@@
 al.
-proposed a new technique named permutation invariant training (PIT) that avoids the label ambiguity problem by treating the reference source streams as a set rather than an ordered list.
+proposed a new technique called permutation invariant training (PIT) that avoids the label ambiguity problem by treating the reference source streams as a set rather than an ordered list.
+One key innovation of PIT is to directly train on all possible permutations of sources ($S!$ possible assignments) and to compute the combined pairwise MSE for every assignment between each reference and the estimated source.
+PIT then directly minimized the MSE of the assignment with the lowest total MSE of every permutation.
+This means that the model is directly trained on label assignment and error evaluation.
+
+PIT operates on a sequence of $N$ input frames termed a \textit{meta-frame} in order to make use of the available contextual information in the sequence.
+It also produces a window of frames as an output and minimizes the separation error at this meta-frame level.
+Since the input is shifted one (or more) frames and a new meta-frame is computed and the assignment is computed for each meta-frame, the output of PIT may produce different speaker assignments for each frame.
+This means that, although the label permutation problem at training time has been resolved, the output of PIT can still be improved by tracing the speaker assignments for each frame as a post-processing step.
+
+PIT results.
+
+uPIT improvements.
+
+uPIT results.
 

### 2018-12-13 paper/main.tex
uPIT results.

@@
 
-PIT results.
+The results of PIT are impressive and exceed 10 dB SDR improvements for the DNN and CNN approaches with optimal speaker assignment.
+The ideal ratio mask (IRM) results indicate the best possible improvement and are only slightly better than PIT with 12.3 dB SDR.
+Optimal speaker assignment means that as a post-processing step the speaker assignment from PIT is improved by tracing the speakers across meta-frames and determining the optimal speaker assignment for each frame.
+This improves performance because the PIT method assumes that the output-speaker does not switch across frames.
+Without the optimal assignment the performance of PIT degrades from a high of 10.9 dB to 5.6 - 7.8 dB.
 
-uPIT improvements.
+The authors of PIT improved their technique by proposing uPIT which is an enhancement to PIT that includes an utterance-level cost function which solves the speaker tracing problem by causing the Recurrent Neural Network (RNN) to align frames from a particular speaker to the same output stream.
+uPIT (like PIT) is also a much simpler model than other proposed deep learning algorithms and the trained networks have strong generalization performance to unseen data.
+In the original PIT algorithm the optimal permutation is computed for each meta-frame which can result in the assignment of speakers to output streams to switch during inference.
+With uPIT the optimal permutation is calculated across the entire utterance assuming the same assignment for all output frames.
+This works in practice by using recurrent neural networks such as deep LSTM (long short-term memory) RNNs and bi-directional LSTM RNNs.
+Since a recurrent neural network has access to data from past frames (and in a BLSTM information both the past and future frames are available) uPIT does not need to be trained on a meta-frame of $N$ stacked frames as PIT did.
 
-uPIT results.
+The results of uPIT are similar to the performance of PIT with optimal speaker tracing.
+Using a BLSTM and a phase sensitive mask (PSM, also referred to as a cIRM - a complex ideal ratio mask) uPIT achieves results near 11 dB SDR.
+The maximum performance from the ideal PSM is about 15 dB SDR (an improvement over the 12.5 dB SDR possible with a standard IRM).
+Based on its performance and (relative) simplicity We chose uPIT as the deep learning solution that we evaluated in our project.
 
@@
 \subsection{CSR-I(WSJ0) Complete}
-For purposes of comparison with existing scholarship, each approach was developed based on the WSJ0 dataset, a corpus of Wall Street Journal text data organized by speakers.
+For purposes of comparison with existing scholarship, each approach was developed based on the WSJ0 dataset, a corpus of read Wall Street Journal text data organized by speakers.
 

### 2018-12-13 paper/main.tex
Cleanup math stuff mostly and some typos.

@@
 
+\usepackage{bm}
+\usepackage{booktabs}
+
 % Other popular packages for formatting tables and equations include:
 
-%\usepackage{array}
+\usepackage{array}
 % Frank Mittelbach's and David Carlisle's array.sty which improves the
@@
 \subsection{NMF, IRM and cIRM}
-Nonnegative matrix factorization (NMF) is used to represent high-dminesional data as the product of two matrices (typically referenced as \textit{W} and \textit{H}).
-In the case of an audio signal, these matrices can be viewed as a representation of the signal's frequency spectrum (\textit{W}) and the corresponding activations (\textit{H}).
+Nonnegative matrix factorization (NMF) is used to represent high-dimensional data as the product of two matrices (typically referenced as $W$ and $H$).
+In the case of an audio signal, these matrices can be viewed as a representation of the signal's frequency spectrum ($W$) and the corresponding activations ($H$).
 
@@
 
-Using the frequency matrices associated with the sample from each speaker, a new set of activations (\textit{W}) was generated from the mixed signal composed of both individual samples.
+Using the frequency matrices associated with the sample from each speaker, a new set of activations ($W$) was generated from the mixed signal composed of both individual samples.
 A magnitude masking matrix, reflecting the magnitudes associated with the frequency representation for each speaker at each time frame, was generated using this new set of activations and the following formula:
 
-\begin{center}
-$W_{S1}H_{(1:30,:)}$
-
-\rule[1pt]{100pt}{.4pt}
-
-$W_{S1}H_{(1:30,:)}+W_{S2}H_{(31:60,:)}  $
-
-\end{center}
+\[
+\dfrac{W_{S1}H_{(1:30,:)}}
+{W_{S1}H_{(1:30,:)}+W_{S2}H_{(31:60,:)}}
+\]
 
@@
 
-Using the STFT of a single speaker audio sample (speaker 1) and the STFT of the mixed audio sample that included speaker 1 [Figure 1], an IRM was generated based on the formula:
-\begin{center}
-$S_{(t,f)}^2$
+Using the STFT of a single speaker audio sample (speaker 1) and the STFT of the mixed audio sample that included speaker 1 (Figure \ref{fig:spec_speaker_1}), an IRM was generated based on the formula:
 
-\rule[1pt]{40pt}{.4pt}
-
-$S_{(t,f)}^2 + N_{(t,f)}^2  $
-
-\end{center}
+\[
+\dfrac{S_{(t,f)}^2}{S_{(t,f)}^2 + N_{(t,f)}^2}
+\]
 
@@
 
-where S is the STFT generated from the isolated speaker 1 audio signal and N is the STFT generated fromm the mixed audio signal that included the same audio sample of speaker 1.\cite{DBLP:journals/corr/abs-1708-07524}
+where $S$ is the STFT generated from the isolated speaker 1 audio signal and $N$ is the STFT generated from the mixed audio signal that included the same audio sample of speaker 1.\cite{DBLP:journals/corr/abs-1708-07524}
 
@@
     \centering  
-     \caption{\label{Fig:Spectrogram of isolated amd mixed audio sample from speaker 1}Spectrogram of isolated amd mixed audio sample from speaker 1}  
+     \caption{\label{fig:spec_speaker_1}Spectrogram of isolated and mixed audio sample from speaker 1}  
     \includegraphics[width=0.4\textwidth]{cIRM_same.png}  
@@
 
-\begin{center}
-$Y_r S_r + Y_i S_i \hspace{1.6cm} Y_r S_i - Y_i S_r$
-
-\rule[3pt]{45pt}{.4pt} \hspace{.3cm} + \hspace{.4cm} \textit{i} \hspace{.1cm} \rule[3pt]{45pt}{.4pt}
-
-$Y_r^2 + Y_i^2  \hspace{2.0cm} Y_r^2 + Y_i^2$
-
-\end{center}
+\[
+\dfrac{Y_r S_r + Y_i S_i}{Y_r^2 + Y_i^2} +
+i\dfrac{Y_r S_i - Y_i S_r}{Y_r^2 + Y_i^2}
+\]
 
 \begin{flushleft}
-where S is the complex STFT representation generated from the isolated speaker 1 audio signal and Y is the complex STFT generated form the mixed audio signal that included the same audio sample of speaker 1.\cite{DBLP:journals/corr/abs-1708-07524}
+where $S$ is the complex STFT representation generated from the isolated speaker 1 audio signal and $Y$ is the complex STFT generated form the mixed audio signal that included the same audio sample of speaker 1.\cite{DBLP:journals/corr/abs-1708-07524}
 \end{flushleft}
@@
 With the NMF method, we were unable to generate a substantial separation between the original isolated signal and the mixed audio.
-Both IMR approaches produced substantially better results with the cIMR method proving to be the best method for this type of separation on a consistent basis [Figure 2].
+Both IMR approaches produced substantially better results with the cIMR method proving to be the best method for this type of separation on a consistent basis [Figure \ref{tab:masking_snr}].
 
@@
 \end{figure}
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{r r r}
+        Label &Waveform & Kurtosis \\
+
+        \midrule
+        0 & s.wav & 7.29 \\
+        1 & x1.wav & 4.91 \\
+        2 & x2.wav & 3.87 \\
+        \midrule
+    \end{tabular}
+    \caption{Table Caption}
+    \label{tab:masking_snr}
+\end{table}
 

### 2018-12-13 paper/main.tex
Document dataset.

@@
 This task is particularly challenging where the input is monophonic, there is no control over environmental conditions, the number of speakers is unknown and there are no samples of the speakers from which to train.
-Using the CSR-I(WSJ0) Complete dataset\cite{WSJintro}, we tested several algorithms for accomplishing monaural speech separation including classification (KMeans), various masking approaches (NMF, Ideal Ratio Mask[IRM], and complex Ideal Ratio Mask [cIRM]), and permutation-based deep learning approaches (Permutation Invariant Training [PIT] and utterance-level Permutation Invariant Training [uPIT]).
+Using the CSR-I(WSJ0) Complete dataset\cite{garofolo:wsj0}, we tested several algorithms for accomplishing monaural speech separation including classification (KMeans), various masking approaches (NMF, Ideal Ratio Mask[IRM], and complex Ideal Ratio Mask [cIRM]), and permutation-based deep learning approaches (Permutation Invariant Training [PIT] and utterance-level Permutation Invariant Training [uPIT]).
 
@@
 The authors found that existing datasets were too limited in size or were not well fitted to their task.
-The deep learning techniques also require a large training set and so the authors adapted the Wall Street Journal (WSJ0) \cite{WSJintro} single-channel speaker corpus to produce a new dataset called WSJ0-2mix \cite{WSJmix} that was used by the later literature for comparison with DPCL.
+The deep learning techniques also require a large training set and so the authors adapted the Wall Street Journal (WSJ0) \cite{garofolo:wsj0} single-channel speaker corpus to produce a new dataset called WSJ0-2mix \cite{WSJmix} that was used by the later literature for comparison with DPCL.
 
@@
 \section{Dataset Preparation}
-\subsection{CSR-I(WSJ0) Complete}
-For purposes of comparison with existing scholarship, each approach was developed based on the WSJ0 dataset, a corpus of read Wall Street Journal text data organized by speakers.
+\subsection{CSR-I (WSJ0) Complete}
+For purposes of comparison with existing scholarship, each approach was developed based on a mixture of audio extracted from the WSJ0 dataset, a corpus of read speech texts drawn from articles in the Wall Street Journal.
+WSJ0 \cite{garofolo:wsj0} consists of over 35,000 audio samples organized into training and evaluation data sets organized by speaker and gender.
+The audio files are stored in a format called Sphere using the Shorten compression algorithm.
+The Linguistic Data Consortium website hosts a small utility called sph2pipe \cite{ldc:sph2pipe} which we used to convert the files from the Sphere format to waveform format.
+We wrote a small script to automate this process and also to merge the separate disks of the WSJ0 dataset into a single folder.
 
 \subsection{WSJ0-mix}
+The WSJ0-mix dataset was developed by the authors of the DPCL \cite{DBLP:journals/corr/HersheyCRW15} paper by mixing selections from the WSJ0 training set \texttt{si\_tr\_s} and mixing them together with different levels of signal-to-noise ratios (SNR) selected from 0 dB to 5dB.
+An additional cross validation was also generated from the training data and an evaluation set was generated with the same process but using the WSJ0 evaluation set \texttt{si\_et\_05} and development set \texttt{si\_dt\_05} which use different speakers than the training set.
+We found the definition of these mixtures and MATLABS scripts \cite{WSJmix} to produce the dataset from the WSJ0 data.
+We made some small adjustments to the scripts so that they would run in Octave, a free, open-source alternative to MATLAB, and to skip the step of generating 16 kHz outputs and the 3-speaker mixtures since we would not be using them.
 
-Developed based on...\cite{WSJmix}.
+The scripts produce 20,000 training files, 5,000 cross validation files, and 3,000 test files.
+Each of these includes the mixed audio as well as the independent sources which have had the correct SNR adjustments made to them.
+We further selected 1,000 training examples and 100 cross-validation and testing examples since we our algorithm was not able to finish running on the full dataset.
 

### 2018-12-13 paper/main.tex
Edits from Jamie.

@@
 
-\subsection{Jamie's Lit Review}
+\subsection{Signal Approximation}
+
+There are several approaches to speech separation based on the general concept that audio can be represented as the combination of a frequency matrix and a corresponding matrix temporal activations.
+Approach the the task as a supervised learning problem,  can be used to create a masking
 
@@
 The audio files are stored in a format called Sphere using the Shorten compression algorithm.
-The Linguistic Data Consortium website hosts a small utility called sph2pipe \cite{ldc:sph2pipe} which we used to convert the files from the Sphere format to waveform format.
+The Linguistic Data Consortium website provides a small utility called sph2pipe \cite{ldc:sph2pipe} which we used to convert the files from the Sphere format to waveform format.
 We wrote a small script to automate this process and also to merge the separate disks of the WSJ0 dataset into a single folder.
@@
     \centering  
-     \caption{\label{Fig:speaker diarization}2 speakers diarization}  
+     \caption{\label{Fig:speaker diarization 1}2 speakers diarization}  
     \includegraphics[width=0.4\textwidth]{kmeans01.png} 
@@
     \centering  
-     \caption{\label{Fig:speaker diarization}2 speakers diarization}  
+     \caption{\label{Fig:speaker diarization 2}2 speakers diarization}  
     \includegraphics[width=0.4\textwidth]{kmeans02.png}
@@
 With the NMF method, we were unable to generate a substantial separation between the original isolated signal and the mixed audio.
-Both IMR approaches produced substantially better results with the cIMR method proving to be the best method for this type of separation on a consistent basis [Figure \ref{tab:masking_snr}].
+Both IMR approaches produced substantially better results with the cIMR method proving to be the best method for this type of separation on a consistent basis [Figure \ref{Fig:masking_snr}].
 
@@
     \centering  
-     \caption{\label{Fig:Comparison of Average SNR for Masking Methods 1}Comparison of Average SNR for Masking Methods}  
-    \includegraphics[width=0.3\textwidth]{Mask_Res.png}  
-\end{figure}
+     \caption{\label{Fig:masking_snr}Average SNR for Masking Methods}
 
-\begin{table}[h]
-    \centering
-    \begin{tabular}{r r r}
-        Label &Waveform & Kurtosis \\
+  \begin{tabular}{r r r}\\
 
+        Method & SNR \\
         \midrule
-        0 & s.wav & 7.29 \\
-        1 & x1.wav & 4.91 \\
-        2 & x2.wav & 3.87 \\
+        NMF & 0.014 \\
+        IRM & 6.380 \\
+        cIRM & 15.070 \\
         \midrule
     \end{tabular}
-    \caption{Table Caption}
-    \label{tab:masking_snr}
-\end{table}
+
+\end{figure}
 

### 2018-12-14 paper/main.tex
Minor updates and PIT diagram.

@@
 This task is particularly challenging where the input is monophonic, there is no control over environmental conditions, the number of speakers is unknown and there are no samples of the speakers from which to train.
-Using the CSR-I(WSJ0) Complete dataset\cite{garofolo:wsj0}, we tested several algorithms for accomplishing monaural speech separation including classification (KMeans), various masking approaches (NMF, Ideal Ratio Mask[IRM], and complex Ideal Ratio Mask [cIRM]), and permutation-based deep learning approaches (Permutation Invariant Training [PIT] and utterance-level Permutation Invariant Training [uPIT]).
+Using the CSR-I(WSJ0) Complete dataset\cite{garofolo:wsj0}, we tested several algorithms for accomplishing monaural speech separation including classification (KMeans), various masking approaches (Nonnegative Matrix Factorization [NMF], Ideal Ratio Mask[IRM], and complex Ideal Ratio Mask [cIRM]), and permutation-based deep learning approaches (Permutation Invariant Training [PIT] and utterance-level Permutation Invariant Training [uPIT]).
 
@@
 
-There are several approaches to speech separation based on the general concept that audio can be represented as the combination of a frequency matrix and a corresponding matrix temporal activations.
-Approach the the task as a supervised learning problem,  can be used to create a masking
+There are several approaches to signal separation (e.g., denoising, diarization) based on the general concept that audio can be represented as the combination of a frequency matrix and a corresponding matrix temporal activations.
+As a supervised learning problem, a time-frequency  representation can be used to emphasize the points at which a target signal dominates the mixed audio and deemphasize the points where an alternate speaker (or other noise source) predominates.[\cite{}]\\
+
+Masking functions range from the Ideal Binary Mask (IBM), in which a masking matrix is used to turn certain frequencies on or off at given moments based on a binary (0/1) assessment of the dominant source at the given time and frequency.[\cite{}] IRM provides a more nuanced approach by representing the relative dominance of each source at a given time and frequency as a ratio of the source signals rather than a binary mask.{\cite{}}
 
@@
 
-\subsection{PIT and uPIT}
+\subsection{uPIT}
 
-..
-.... Add text here....
+\begin{figure}[h!]
+    \centering  
+     \caption{\label{fig:pit_diagram}PIT 2 speaker separation model}  
+    \includegraphics[width=0.4\textwidth]{PIT.png}  
+\end{figure}
 

### 2018-12-14 paper/main.tex
Number equations and PIT detail.

@@
 To perform NMF, we transformed isolated audio samples of two speakers (speaker 1 and speaker 2) to a frequency-time representation using short-time Fourier transform (STFT) before decomposing each signal using the following iterative update rules:
-\newline
 
-$W = W \bigodot \frac{\frac{X}{WH}H^T}{1{FxT}H^T}$ \hspace{1cm} $H = H \bigodot \frac{W^T\frac{X}{WH}}{W^T1^{FxT}}$
-\newline
+\begin{equation}
+\begin{array}{ll}
+W = W \bigodot \frac{\frac{X}{WH}H^T}{1{FxT}H^T} &
+H = H \bigodot \frac{W^T\frac{X}{WH}}{W^T1^{FxT}}
+\end{array}
+\end{equation}
 
@@
 
-\[
+\begin{equation}
 \dfrac{W_{S1}H_{(1:30,:)}}
 {W_{S1}H_{(1:30,:)}+W_{S2}H_{(31:60,:)}}
-\]
+\end{equation}
 
@@
 
-\[
+\begin{equation}
 \dfrac{S_{(t,f)}^2}{S_{(t,f)}^2 + N_{(t,f)}^2}
-\]
+\end{equation}
 
@@
 
-\[
+\begin{equation}
 \dfrac{Y_r S_r + Y_i S_i}{Y_r^2 + Y_i^2} +
 i\dfrac{Y_r S_i - Y_i S_r}{Y_r^2 + Y_i^2}
-\]
+\end{equation}
 
@@
 
-\subsection{uPIT}
+\subsection{PIT and uPIT}
+
+Utterance-level Permutation Invariant Training (uPIT) is an advancement on the Permutation Invariant Training model depicted in Figure \ref{fig:pit_diagram} where the reference signals are provided to the training model as a set rather than a list by calculating the error on all permutations of the output signals and then optimizing the model on the permutation with the lowest mean squared error (MSE).
+A traditional approach is to minimize the MSE between the estimated masks and the ideal masks such as in
+
+\begin{equation}
+J_m = \dfrac{1}{B} \sum_{s=1}^S || \Hat{\mathbf{M}}_S - \mathbf{M}_S ||^2_F
+\end{equation}
+
+where $B = T \times N \times S$ is the count of all time-frequency units over all sources $S$ and $F$ refers to the Frobenius norm.
+However, to avoid issues predicting masks in silence segments and to directly calculate the error relative to the signal reconstruction, PIT and uPIT minimize the MSE between the actual signals by applying the mask to the original spectra and then calculating the error.
+For PIT using the PSM with complex values the cost function is
+
+\begin{equation}
+J_{psm} = \dfrac{1}{B} \sum_{s=1}^S || \Hat{\mathbf{M}}_S \odot \mathbf{R} - \mathbf{A}_S \odot \cos(\theta_y - \theta_s) ||^2_F  
+\end{equation}
 

### 2018-12-14 paper/main.tex
Cleaning up and finishing PIT.

@@
 
+\DeclareMathOperator*{\argmax}{arg\ max}
+\DeclareMathOperator*{\argmin}{arg\ min}
+
 % Your document starts here!
@@
 % Define document title and author
-	\title{E511-Signal Processing-Final Project}
+	\title{Monaural Single-Source Speaker Separation \\
+	{\large E511 - Signal Processing - Final Project}}
 	\author{Jonathan Branam, Jamie Israel, Jinju(Hellen) Jiang}
@@
 
+\begin{figure}[h]
+    \centering  
+     \caption{\label{fig:pit_diagram}PIT 2 speaker separation model}  
+    \includegraphics[width=0.4\textwidth]{PIT.png}  
+\end{figure}
+
 Utterance-level Permutation Invariant Training (uPIT) is an advancement on the Permutation Invariant Training model depicted in Figure \ref{fig:pit_diagram} where the reference signals are provided to the training model as a set rather than a list by calculating the error on all permutations of the output signals and then optimizing the model on the permutation with the lowest mean squared error (MSE).
@@
 However, to avoid issues predicting masks in silence segments and to directly calculate the error relative to the signal reconstruction, PIT and uPIT minimize the MSE between the actual signals by applying the mask to the original spectra and then calculating the error.
-For PIT using the PSM with complex values the cost function is
+For PIT using PSM with complex values the cost function is
 
 \begin{equation}
-J_{psm} = \dfrac{1}{B} \sum_{s=1}^S || \Hat{\mathbf{M}}_S \odot \mathbf{R} - \mathbf{A}_S \odot \cos(\theta_y - \theta_s) ||^2_F  
+J_{\phi^*} = \dfrac{1}{B} \sum_{s=1}^S || \Hat{\mathbf{M}}_s \odot \mathbf{R} - 
+\mathbf{A}_{\phi^*(s)} \odot \cos(\theta_y - \theta_{\phi^*(s)}) ||^2_F  
 \end{equation}
 
-\begin{figure}[h!]
-    \centering  
-     \caption{\label{fig:pit_diagram}PIT 2 speaker separation model}  
-    \includegraphics[width=0.4\textwidth]{PIT.png}  
-\end{figure}
+where $\phi^*$ is the permutation with the lowest MSE:
+\begin{equation}
+\phi^* = \argmin_{\phi \in \mathcal{P}} \dfrac{1}{B} \sum_{s=1}^S || \Hat{\mathbf{M}}_s \odot \mathbf{R} - 
+\mathbf{A}_{\phi(s)} \odot \cos(\theta_y - \theta_{\phi(s)}) ||^2_F  
+\end{equation}
+
+The major difference in uPIT is instead of calculating the permutation error for every meta-frame uPIT calculates the error \textit{across the entire utterance}.
+uPIT relies on using recurrent neural networks (RNNs) to provide the context necessary for good separation and for selecting the best permutation across the entire utterance.
+The cost function for uPIT is the same as for PIT but is calculated for the entire utterance rather than for each meta-frame.
 
@@
 
+\subsection{uPIT}
+
+Our implementation of uPIT uses a BLSTM with 3 hidden layers of size 896.
+We used the Adam optimizer and set a $0.5$ rate for dropout.
+We trained using Google Colab free instances which allow using a Tesla K80 GPU with 12 Gb or RAM.
+We found that training with 1,000 samples for 60 epochs took about 5 hours in this setup.
+The Google Colab GPU will disconnect after 12 hours so we were unable to complete a full training run.
+We experimented with saving off the model parameters and restarting the training once the 12 hours had expired but were unable to complete the implementation with enough time to train again.
+
 \section{Conclusion}

### 2018-12-14 paper/main.tex
Some small updates.

@@
 There are several approaches to signal separation (e.g., denoising, diarization) based on the general concept that audio can be represented as the combination of a frequency matrix and a corresponding matrix temporal activations.
-As a supervised learning problem, a time-frequency  representation can be used to emphasize the points at which a target signal dominates the mixed audio and deemphasize the points where an alternate speaker (or other noise source) predominates.[\cite{}]\\
+As a supervised learning problem, a time-frequency  representation can be used to emphasize the points at which a target signal dominates the mixed audio and deemphasize the points where an alternate speaker (or other noise source) predominates.\cite{DBLP:journals/corr/abs-1708-07524}\\
 
-Masking functions range from the Ideal Binary Mask (IBM), in which a masking matrix is used to turn certain frequencies on or off at given moments based on a binary (0/1) assessment of the dominant source at the given time and frequency.[\cite{}] IRM provides a more nuanced approach by representing the relative dominance of each source at a given time and frequency as a ratio of the source signals rather than a binary mask.{\cite{}}
+Masking functions range from the Ideal Binary Mask (IBM), in which a masking matrix is used to turn certain frequencies on or off at given moments based on a binary (0/1) assessment of the dominant source at the given time and frequency.\cite{DBLP:journals/corr/abs-1708-07524} IRM provides a more nuanced approach by representing the relative dominance of each source at a given time and frequency as a ratio of the source signals rather than a binary mask.\cite{DBLP:journals/corr/abs-1708-07524}
 
@@
 Since we knew that the audio from call center is the conversation between 2 speakers(customer service and customer), we set k=2 for kmeans clustering.\\
-here it is the results for speaker diarization.
+The results for kmeans speaker diarization are shown in \ref{Fig:speaker diarization 1} and \ref{Fig:speaker diarization 2}.
 \begin{figure}[h!]
@@
 
+Our results with uPIT were not very good, probably due to the small training size that we used.
+We were able to find two open source implementations of uPIT one of them we used as a reference only [CITE 1] and the other [CITE 2] we were able to actually implement in order to verify our results.
+The author was able to achieve results comparable to those published in the uPIT paper and he agreed that we should not expect to match the paper's performance unless we are able to train on the full dataset.
+
 \section{Conclusion}

### 2018-12-14 paper/main.tex

@@
 There are several approaches to signal separation (e.g., denoising, diarization) based on the general concept that audio can be represented as the combination of a frequency matrix and a corresponding matrix temporal activations.
-As a supervised learning problem, a time-frequency  representation can be used to emphasize the points at which a target signal dominates the mixed audio and deemphasize the points where an alternate speaker (or other noise source) predominates.\cite{DBLP:journals/corr/abs-1708-07524}\\
+As a supervised learning problem, a time-frequency  representation can be used to emphasize the points at which a target signal dominates the mixed audio and deemphasize the points where an alternate speaker (or other noise source) predominates.\cite{DBLP:journals/corr/abs-1708-07524}
 
-Masking functions range from the Ideal Binary Mask (IBM), in which a masking matrix is used to turn certain frequencies on or off at given moments based on a binary (0/1) assessment of the dominant source at the given time and frequency.\cite{DBLP:journals/corr/abs-1708-07524} IRM provides a more nuanced approach by representing the relative dominance of each source at a given time and frequency as a ratio of the source signals rather than a binary mask.\cite{DBLP:journals/corr/abs-1708-07524}
+Masking functions range from the Ideal Binary Mask (IBM), in which a masking matrix is used to turn certain frequencies on or off at given moments based on a binary (0/1) assessment of the dominant source at the given time and frequency.\cite{DBLP:journals/corr/abs-1708-07524} IRM provides a more nuanced approach by representing the relative dominance of each source at a given time and frequency as a ratio of the source signals rather than a binary mask.
+And, cIRM is an Ideal Ratio Mask calculated in the complex domain, which allows for a full reconstruction of the audio signal.
+\cite{DBLP:journals/corr/abs-1708-07524}
+
+Each of the aforementioned masking models relies on a global representation of the frequency spectrum associated with a given audio sample.
+More recently, probabilistic approaches have been used to represent the signal as a collection of frequency spectrum (hidden states) over time.
+On such approach, Nonnegative Hidden Markov Model (N-HMM), models a magnitude spectrogram for each latent component (source) at a given time frame as one of many states; like a traditional HMM, movement between states - and the spectral representation governing each latent component - is learned and defined by a transition matrix.
+\cite{10.1007/978-3-642-15995-4_18}
 
@@
 Since we knew that the audio from call center is the conversation between 2 speakers(customer service and customer), we set k=2 for kmeans clustering.\\
-The results for kmeans speaker diarization are shown in \ref{Fig:speaker diarization 1} and \ref{Fig:speaker diarization 2}.
+The results for kmeans speaker diarization are shown in Figure\ref{Fig:speaker diarization 1} and Figure \ref{Fig:speaker diarization 2}.
 \begin{figure}[h!]
@@
 \end{figure}
-\subsubsection{Evaluation\cite{Segmentation}}Diarization Error Rate (DER) is used for evaluation of automatic speech recognition system.Contribution to DER comes from three factors namely, Missed speech rate (MSR), False alarm speech rate (FASR) and Speaker Error.
-When a speech is labeled as non-speech then that error comes under MSR.
-FASR is when a non- speech is detected as a speech segment.
-Speaker error is contributed due to speaker clustering and segmentation.
-This kind of error can be caused if a speaker change is not detected, oversegmentation, erroneously clustered.
-Sum of all three errors contribute to the DER.\\
-In this case, we don't have any labeled data to calculate the training error, we just simply used human ears to justify whether the diarization was good or not.
+\subsubsection{Evaluation\cite{Segmentation}}The common metric to measure how good/bad for speaker diarization experiments is the Diarization Error Rate(DER), it contains 3 main parts:
+\begin{itemize}
+    \item The first one is called Missed Speech Rate(MSR), whichi is percentage of scored time that a hypothesized non-speech segment corresponds to a reference speaker segment.
+    \item The second one is False alarm speech rate (FASR), which is percentage of scored time that a hypothesized speaker is labelled as a non-speech in the reference.
+    \item The third one is Speaker error rate(SER), which is percentage of scored time that a speaker ID is assigned to the wrong speaker.
+    \item The last one is Overlap speaker rate(OSR), which is percentage of scored time that some of the multiple speakers in a segment do not get assigned to any speaker.
+\end{itemize}
+
+So the Diarization Error Rate(DER) can be calculated as the sum of the above 4 rates(MSR+FASR+SER+OSR)\\
+In this case, we don't have any labeled data to identify which segmentation is speaker 1/2 to calculate the training error, we just simply used human ears to justify whether the diarization was good or not.
 For some training data, it works well, but some training data, it performed really bad.
@@
 \section{Conclusion}
-...
+We got our best results from the cIRM with NMF separation. uPIT needed more training time and Kmeans...
+
 \newline 

### 2018-12-14 proposal/sig-alternate.tex

% This is "sig-alternate.tex" V2.0 May 2012
% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
%
% This example file demonstrates the use of the 'sig-alternate.cls'
% V2.5 LaTeX2e document class file. It is for those submitting
% articles to ACM Conference Proceedings WHO DO NOT WISH TO
% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
% The 'sig-alternate.cls' file will produce a similar-looking,
% albeit, 'tighter' paper resulting in, invariably, fewer pages.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V2.5) produces:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) NO page numbers
%
% as against the acm_proc_article-sp.cls file which
% DOES NOT produce 1) thru' 3) above.
%
% Using 'sig-alternate.cls' you have control, however, from within
% the source .tex file, over both the CopyrightYear
% (defaulted to 200X) and the ACM Copyright Data
% (defaulted to X-XXXXX-XX-X/XX/XX).
% e.g.
% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
%
% ---------------------------------------------------------------------------------------------------------------
% This .tex source is an example which *does* use
% the .bib file (from which the .bbl file % is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% ================= IF YOU HAVE QUESTIONS =======================
% Questions regarding the SIGS styles, SIGS policies and
% procedures, Conferences etc. should be sent to
% Adrienne Griscti (griscti@acm.org)
%
% Technical questions _only_ to
% Gerald Murray (murray@hq.acm.org)
% ===============================================================
%
% For tracking purposes - this is V2.0 - May 2012

\documentclass{sig-alternate}

\begin{document}
%
% --- Author Metadata here ---
\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{Alternate {\ttlit ACM} SIG Proceedings Paper in LaTeX
Format\titlenote{(Produces the permission block, and
copyright information). For use with
SIG-ALTERNATE.CLS. Supported by ACM.}}
\subtitle{[Extended Abstract]
\titlenote{A full version of this paper is available as
\textit{Author's Guide to Preparing ACM SIG Proceedings Using
\LaTeX$2_\epsilon$\ and BibTeX} at
\texttt{www.acm.org/eaddress.htm}}}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{8} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
       \affaddr{Institute for Clarity in Documentation}\\
       \affaddr{1932 Wallamaloo Lane}\\
       \affaddr{Wallamaloo, New Zealand}\\
       \email{trovato@corporation.com}
% 2nd. author
\alignauthor
G.K.M. Tobin\titlenote{The secretary disavows
any knowledge of this author's actions.}\\
       \affaddr{Institute for Clarity in Documentation}\\
       \affaddr{P.O. Box 1212}\\
       \affaddr{Dublin, Ohio 43017-6221}\\
       \email{webmaster@marysville-ohio.com}
% 3rd. author
\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
one who did all the really hard work.}\\
       \affaddr{The Th{\o}rv{\"a}ld Group}\\
       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
       \affaddr{Hekla, Iceland}\\
       \email{larst@affiliation.org}
\and  % use '\and' if you need 'another row' of author names
% 4th. author
\alignauthor Lawrence P. Leipuner\\
       \affaddr{Brookhaven Laboratories}\\
       \affaddr{Brookhaven National Lab}\\
       \affaddr{P.O. Box 5000}\\
       \email{lleipuner@researchlabs.org}
% 5th. author
\alignauthor Sean Fogarty\\
       \affaddr{NASA Ames Research Center}\\
       \affaddr{Moffett Field}\\
       \affaddr{California 94035}\\
       \email{fogartys@amesres.org}
% 6th. author
\alignauthor Charles Palmer\\
       \affaddr{Palmer Research Laboratories}\\
       \affaddr{8600 Datapoint Drive}\\
       \affaddr{San Antonio, Texas 78229}\\
       \email{cpalmer@prl.com}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle
\begin{abstract}
This paper provides a sample of a \LaTeX\ document which conforms,
somewhat loosely, to the formatting guidelines for
ACM SIG Proceedings. It is an {\em alternate} style which produces
a {\em tighter-looking} paper and was designed in response to
concerns expressed, by authors, over page-budgets.
It complements the document \textit{Author's (Alternate) Guide to
Preparing ACM SIG Proceedings Using \LaTeX$2_\epsilon$\ and Bib\TeX}.
This source file has been written with the intention of being
compiled under \LaTeX$2_\epsilon$\ and BibTeX.

The developers have tried to include every imaginable sort
of ``bells and whistles", such as a subtitle, footnotes on
title, subtitle and authors, as well as in the text, and
every optional component (e.g. Acknowledgments, Additional
Authors, Appendices), not to mention examples of
equations, theorems, tables and figures.

To make best use of this sample document, run it through \LaTeX\
and BibTeX, and compare this source code with the printed
output produced by the dvi file. A compiled PDF version
is available on the web page to help you with the
`look and feel'.
\end{abstract}

% A category with the (minimum) three required fields
\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

\terms{Theory}

\keywords{ACM proceedings, \LaTeX, text tagging}

\section{Introduction}
The \textit{proceedings} are the records of a conference.
ACM seeks to give these conference by-products a uniform,
high-quality appearance.  To do this, ACM has some rigid
requirements for the format of the proceedings documents: there
is a specified format (balanced  double columns), a specified
set of fonts (Arial or Helvetica and Times Roman) in
certain specified sizes (for instance, 9 point for body copy),
a specified live area (18 $\times$ 23.5 cm [7" $\times$ 9.25"]) centered on
the page, specified size of margins (1.9 cm [0.75"]) top, (2.54 cm [1"]) bottom
and (1.9 cm [.75"]) left and right; specified column width
(8.45 cm [3.33"]) and gutter size (.83 cm [.33"]).

The good news is, with only a handful of manual
settings\footnote{Two of these, the {\texttt{\char'134 numberofauthors}}
and {\texttt{\char'134 alignauthor}} commands, you have
already used; another, {\texttt{\char'134 balancecolumns}}, will
be used in your very last run of \LaTeX\ to ensure
balanced column heights on the last page.}, the \LaTeX\ document
class file handles all of this for you.

The remainder of this document is concerned with showing, in
the context of an ``actual'' document, the \LaTeX\ commands
specifically available for denoting the structure of a
proceedings paper, rather than with giving rigorous descriptions
or explanations of such commands.

\section{The {\secit Body} of The Paper}
Typically, the body of a paper is organized
into a hierarchical structure, with numbered or unnumbered
headings for sections, subsections, sub-subsections, and even
smaller sections.  The command \texttt{{\char'134}section} that
precedes this paragraph is part of such a
hierarchy.\footnote{This is the second footnote.  It
starts a series of three footnotes that add nothing
informational, but just give an idea of how footnotes work
and look. It is a wordy one, just so you see
how a longish one plays out.} \LaTeX\ handles the numbering
and placement of these headings for you, when you use
the appropriate heading commands around the titles
of the headings.  If you want a sub-subsection or
smaller part to be unnumbered in your output, simply append an
asterisk to the command name.  Examples of both
numbered and unnumbered headings will appear throughout the
balance of this sample document.

Because the entire article is contained in
the \textbf{document} environment, you can indicate the
start of a new paragraph with a blank line in your
input file; that is why this sentence forms a separate paragraph.

\subsection{Type Changes and {\subsecit Special} Characters}
We have already seen several typeface changes in this sample.  You
can indicate italicized words or phrases in your text with
the command \texttt{{\char'134}textit}; emboldening with the
command \texttt{{\char'134}textbf}
and typewriter-style (for instance, for computer code) with
\texttt{{\char'134}texttt}.  But remember, you do not
have to indicate typestyle changes when such changes are
part of the \textit{structural} elements of your
article; for instance, the heading of this subsection will
be in a sans serif\footnote{A third footnote, here.
Let's make this a rather short one to
see how it looks.} typeface, but that is handled by the
document class file. Take care with the use
of\footnote{A fourth, and last, footnote.}
the curly braces in typeface changes; they mark
the beginning and end of
the text that is to be in the different typeface.

You can use whatever symbols, accented characters, or
non-English characters you need anywhere in your document;
you can find a complete list of what is
available in the \textit{\LaTeX\
User's Guide}\cite{Lamport:LaTeX}.

\subsection{Math Equations}
You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of
the three are discussed in the next sections.

\subsubsection{Inline (In-text) Equations}
A formula that appears in the running text is called an
inline or in-text formula.  It is produced by the
\textbf{math} environment, which can be
invoked with the usual \texttt{{\char'134}begin. . .{\char'134}end}
construction or with the short form \texttt{\$. . .\$}. You
can use any of the symbols and structures,
from $\alpha$ to $\omega$, available in
\LaTeX\cite{Lamport:LaTeX}; this section will simply show a
few examples of in-text equations in context. Notice how
this equation: \begin{math}\lim_{n\rightarrow \infty}x=0\end{math},
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsubsection{Display Equations}
A numbered display equation -- one set off by vertical space
from the text and centered horizontally -- is produced
by the \textbf{equation} environment. An unnumbered display
equation is produced by the \textbf{displaymath} environment.

Again, in either environment, you can use any of the symbols
and structures available in \LaTeX; this section will just
give a couple of examples of display equations in context.
First, consider the equation, shown as an inline equation above:
\begin{equation}\lim_{n\rightarrow \infty}x=0\end{equation}
Notice how it is formatted somewhat differently in
the \textbf{displaymath}
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}\sum_{i=0}^{\infty} x + 1\end{displaymath}
and follow it with another numbered equation:
\begin{equation}\sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f\end{equation}
just to demonstrate \LaTeX's able handling of numbering.

\subsection{Citations}
Citations to articles \cite{bowman:reasoning,
clark:pct, braams:babel, herlihy:methodology},
conference proceedings \cite{clark:pct} or
books \cite{salas:calculus, Lamport:LaTeX} listed
in the Bibliography section of your
article will occur throughout the text of your article.
You should use BibTeX to automatically produce this bibliography;
you simply need to insert one of several citation commands with
a key of the item cited in the proper location in
the \texttt{.tex} file \cite{Lamport:LaTeX}.
The key is a short reference you invent to uniquely
identify each work; in this sample document, the key is
the first author's surname and a
word from the title.  This identifying key is included
with each item in the \texttt{.bib} file for your article.

The details of the construction of the \texttt{.bib} file
are beyond the scope of this sample document, but more
information can be found in the \textit{Author's Guide},
and exhaustive details in the \textit{\LaTeX\ User's
Guide}\cite{Lamport:LaTeX}.

This article shows only the plainest form
of the citation command, using \texttt{{\char'134}cite}.
This is what is stipulated in the SIGS style specifications.
No other citation format is endorsed or supported.

\subsection{Tables}
Because tables cannot be split across pages, the best
placement for them is typically the top of the page
nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \textbf{table} to enclose the table's contents and
the table caption.  The contents of the table itself must go
in the \textbf{tabular} environment, to
be aligned properly in rows and columns, with the desired
horizontal and vertical rules.  Again, detailed instructions
on \textbf{tabular} material
is found in the \textit{\LaTeX\ User's Guide}.

Immediately following this sentence is the point at which
Table 1 is included in the input file; compare the
placement of the table here with the table in the printed
dvi output of this document.

\begin{table}
\centering
\caption{Frequency of Special Characters}
\begin{tabular}{|c|c|l|} \hline
Non-English or Math&Frequency&Comments\\ \hline
\O & 1 in 1,000& For Swedish names\\ \hline
$\pi$ & 1 in 5& Common in math\\ \hline
\$ & 4 in 5 & Used in business\\ \hline
$\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
\hline\end{tabular}
\end{table}

To set a wider table, which takes up the whole width of
the page's live area, use the environment
\textbf{table*} to enclose the table's contents and
the table caption.  As with a single-column table, this wide
table will ``float" to a location deemed more desirable.
Immediately following this sentence is the point at which
Table 2 is included in the input file; again, it is
instructive to compare the placement of the
table here with the table in the printed dvi
output of this document.

\begin{table*}
\centering
\caption{Some Typical Commands}
\begin{tabular}{|c|c|l|} \hline
Command&A Number&Comments\\ \hline
\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
\texttt{{\char'134}table}& 300 & For tables\\ \hline
\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
\end{table*}
% end the environment with {table*}, NOTE not {table}!

\subsection{Figures}
Like tables, figures cannot be split across pages; the
best placement for them
is typically the top or the bottom of the page nearest
their initial cite.  To ensure this proper ``floating'' placement
of figures, use the environment
\textbf{figure} to enclose the figure and its caption.

This sample document contains examples of \textbf{.eps}
and \textbf{.ps} files to be displayable with \LaTeX.  More
details on each of these is found in the \textit{Author's Guide}.

\begin{figure}
\centering
%\epsfig{file=fly.eps}
\caption{A sample black and white graphic (.eps format).}
\end{figure}

\begin{figure}
\centering
%\epsfig{file=fly.eps, height=1in, width=1in}
\caption{A sample black and white graphic (.eps format)
that has been resized with the \texttt{epsfig} command.}
\end{figure}

As was the case with tables, you may want a figure
that spans two columns.  To do this, and still to
ensure proper ``floating'' placement of tables, use the environment
\textbf{figure*} to enclose the figure and its caption.
and don't forget to end the environment with
{figure*}, not {figure}!

\begin{figure*}
\centering
%\epsfig{file=flies.eps}
\caption{A sample black and white graphic (.eps format)
that needs to span two columns of text.}
\end{figure*}

Note that either {\textbf{.ps}} or {\textbf{.eps}} formats are
used; use
the \texttt{{\char'134}epsfig} or \texttt{{\char'134}psfig}
commands as appropriate for the different file types.

\begin{figure}
\centering
% \psfig{file=rosette.ps, height=1in, width=1in,}
\caption{A sample black and white graphic (.ps format) that has
been resized with the \texttt{psfig} command.}
\vskip -6pt
\end{figure}

\subsection{Theorem-like Constructs}
Other common constructs that may occur in your article are
the forms for logical constructs like theorems, axioms,
corollaries and proofs.  There are
two forms, one produced by the
command \texttt{{\char'134}newtheorem} and the
other by the command \texttt{{\char'134}newdef}; perhaps
the clearest and easiest way to distinguish them is
to compare the two in the output of this sample document:

This uses the \textbf{theorem} environment, created by
the\linebreak\texttt{{\char'134}newtheorem} command:
\newtheorem{theorem}{Theorem}
\begin{theorem}
Let $f$ be continuous on $[a,b]$.  If $G$ is
an antiderivative for $f$ on $[a,b]$, then
\begin{displaymath}\int^b_af(t)dt = G(b) - G(a).\end{displaymath}
\end{theorem}

The other uses the \textbf{definition} environment, created
by the \texttt{{\char'134}newdef} command:
\newdef{definition}{Definition}
\begin{definition}
If $z$ is irrational, then by $e^z$ we mean the
unique number which has
logarithm $z$: \begin{displaymath}{\log e^z = z}\end{displaymath}
\end{definition}

Two lists of constructs that use one of these
forms is given in the
\textit{Author's  Guidelines}.

There is one other similar construct environment, which is
already set up
for you; i.e. you must \textit{not} use
a \texttt{{\char'134}newdef} command to
create it: the \textbf{proof} environment.  Here
is a example of its use:
\begin{proof}
Suppose on the contrary there exists a real number $L$ such that
\begin{displaymath}
\lim_{x\rightarrow\infty} \frac{f(x)}{g(x)} = L.
\end{displaymath}
Then
\begin{displaymath}
l=\lim_{x\rightarrow c} f(x)
= \lim_{x\rightarrow c}
\left[ g{x} \cdot \frac{f(x)}{g(x)} \right ]
= \lim_{x\rightarrow c} g(x) \cdot \lim_{x\rightarrow c}
\frac{f(x)}{g(x)} = 0\cdot L = 0,
\end{displaymath}
which contradicts our assumption that $l\neq 0$.
\end{proof}

Complete rules about using these environments and using the
two different creation commands are in the
\textit{Author's Guide}; please consult it for more
detailed instructions.  If you need to use another construct,
not listed therein, which you want to have the same
formatting as the Theorem
or the Definition\cite{salas:calculus} shown above,
use the \texttt{{\char'134}newtheorem} or the
\texttt{{\char'134}newdef} command,
respectively, to create it.

\subsection*{A {\secit Caveat} for the \TeX\ Expert}
Because you have just been given permission to
use the \texttt{{\char'134}newdef} command to create a
new form, you might think you can
use \TeX's \texttt{{\char'134}def} to create a
new command: \textit{Please refrain from doing this!}
Remember that your \LaTeX\ source code is primarily intended
to create camera-ready copy, but may be converted
to other forms -- e.g. HTML. If you inadvertently omit
some or all of the \texttt{{\char'134}def}s recompilation will
be, to say the least, problematic.

\section{Conclusions}
This paragraph will end the body of this sample document.
Remember that you might still have Acknowledgments or
Appendices; brief samples of these
follow.  There is still the Bibliography to deal with; and
we will make a disclaimer about that here: with the exception
of the reference to the \LaTeX\ book, the citations in
this paper are to articles which have nothing to
do with the present subject and are used as
examples only.
%\end{document}  % This is where a 'short' article might terminate

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}
This section is optional; it is a location for you
to acknowledge grants, funding, editing assistance and
what have you.  In the present case, for example, the
authors would like to thank Gerald Murray of ACM for
his help in codifying this \textit{Author's Guide}
and the \textbf{.cls} and \textbf{.tex} files that it describes.

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns
\appendix
%Appendix A
\section{Headings in Appendices}
The rules about hierarchical headings discussed above for
the body of the article are different in the appendices.
In the \textbf{appendix} environment, the command
\textbf{section} is used to
indicate the start of each Appendix, with alphabetic order
designation (i.e. the first is A, the second B, etc.) and
a title (if you include one).  So, if you need
hierarchical structure
\textit{within} an Appendix, start with \textbf{subsection} as the
highest level. Here is an outline of the body of this
document in Appendix-appropriate form:
\subsection{Introduction}
\subsection{The Body of the Paper}
\subsubsection{Type Changes and  Special Characters}
\subsubsection{Math Equations}
\paragraph{Inline (In-text) Equations}
\paragraph{Display Equations}
\subsubsection{Citations}
\subsubsection{Tables}
\subsubsection{Figures}
\subsubsection{Theorem-like Constructs}
\subsubsection*{A Caveat for the \TeX\ Expert}
\subsection{Conclusions}
\subsection{Acknowledgments}
\subsection{Additional Authors}
This section is inserted by \LaTeX; you do not insert it.
You just add the names and information in the
\texttt{{\char'134}additionalauthors} command at the start
of the document.
\subsection{References}
Generated by bibtex from your ~.bib file.  Run latex,
then bibtex, then latex twice (to resolve references)
to create the ~.bbl file.  Insert that ~.bbl file into
the .tex source file and comment out
the command \texttt{{\char'134}thebibliography}.
% This next section command marks the start of
% Appendix B, and does not continue the present hierarchy
\section{More Help for the Hardy}
The sig-alternate.cls file itself is chock-full of succinct
and helpful comments.  If you consider yourself a moderately
experienced to expert user of \LaTeX, you may find reading
it useful but please remember not to change it.
%\balancecolumns % GM June 2007
% That's all folks!
\end{document}

### 2018-12-14 paper/main.tex
Updates mostly from Hellen.

@@
 \begin{itemize}
-    \item A chroma vector (Wikipedia) (FMP, p.
-123) is a typically a 12-element feature vector indicating how much energy of each pitch class, {C, C\#, D, D\#, E, ..., B}, is present in the signal.
+    \item A chroma vector (Wikipedia) that consists of 12 feature vectors indicating how much energy of each pitch class,In the signal we used the same notation as used i Western music notation, to present every pitch class.
 Then use mean of each element as one of input for clustering.
-    \item The mel frequency cepstral coefficients (MFCCs) of a signal are a small set of features  which concisely describe the overall shape of a spectral envelope.
-In this case, mfcc computed 13 MFCCs.The very first MFCC, the 0th coefficient, does not convey information relevant to the overall shape of the spectrum.
-It only conveys a constant offset, i.e.
-adding a constant value to the entire spectrum.
-Therefore, here I discard the first MFCC, then I scale the MFCCs such that each coefficient dimension has zero mean and unit variance.Then use mean of MFCCS as one part of input for clustering.
+    \item MFCCs[Wikipedia] are commonly used as features in speech recognition systems, such as the systems which can automatically recognize numbers spoken into a telephone.
+In this experiment, mfcc computed 13 MFCCs.The first MFCC does not represent information related to the overall shape of the spectrum.
+It only contains a constant offset.
+Therefore, here I discard the first MFCC, then I scale the MFCCs such that each coefficient dimension has zero mean and unit variance.Then use mean of MFCCs as one part of input for clustering.
     \item We divide the audio signal into smaller frames.
@@
 \end{itemize}
-\subsubsection{Speaker segmentation}\cite{Segmentation} Speaker segmentation which is also known as acoustic change detection aims to detect speaker change such that each contiguous segment corresponds to single speaker only.
-To find if the two segments correspond to same speaker we have to define some notion of distance metric.
-In this case, I skipped this step and used the features extrated from above steps as input for kmeans clustering.
+\subsubsection{Speaker segmentation}\cite{Segmentation} Speaker segmentation is to split an audio stream into smaller segments, so as every segment ideally contains only one speaker.
+So the key point for speech segment is to find out the change point where speakers changed at that point.
+For example, in our training data, we have conversation between customer service(speaker 0) and customer(speak 1), so we need to find out each change point speaker 0 to speaker 1 or vice versa.
+In this case, I skipped this step and used the features extracted from above steps as input for kmeans clustering.
 \subsubsection{Kmeans Clustering}
@@
         \midrule
-        NMF & 0.014 \\
+        NMF & 1.924 \\
         IRM & 6.380 \\
@@
 
-\newline 
 % Now we need a bibliography:

### 2018-12-15 paper/main.tex
Title update.

@@
 	\title{Monaural Single-Source Speaker Separation \\
-	{\large E511 - Signal Processing - Final Project}}
-	\author{Jonathan Branam, Jamie Israel, Jinju(Hellen) Jiang}
-	\date{December, 13 2018}
+	{\large E511 - Machine Learning for Signal Processing - Final Project}}
+	\author{Jonathan Branam, Jamie Israel, Jinju (Hellen) Jiang}
+	\date{December, 14 2018}
 	\maketitle
@@
     \centering  
-     \caption{\label{Fig:speaker diarization 1}2 speakers diarization}  
+     \caption{\label{Fig:speaker diarization 1}2 speakers diarization output}  
     \includegraphics[width=0.4\textwidth]{kmeans01.png} 
@@
     \centering  
-     \caption{\label{Fig:speaker diarization 2}2 speakers diarization}  
+     \caption{\label{Fig:speaker diarization 2}2 speakers diarization on spectrogram}  
     \includegraphics[width=0.4\textwidth]{kmeans02.png}

### 2018-12-15 code/train.py
uPIT implementation.

@@ -1,0 +1,484 @@
+#!/usr/bin/env python3
+# train.py
+"""Train a uPIT BLSTM
+
+Usage:
+    train.py [options]
+
+Options:
+    -d, --debug                         # output extra debug information
+    --debug-level LEVEL                 # debug level to LEVEL [default: 1]
+                                        #
+    --train-root-dir TRAIN-ROOT-DIR     # training set root directory [default: data/2spk8kmax/tr]
+    --valid-root-dir VALID-ROOT-DIR     # validation set root directory [default: data/2spk8kmax/cv]
+    --mix-dir MIX-DIR                   # directory beneath root with mixtures [default: mix]
+    --src1-dir SRC1-DIR                 # directory beneath root with source 1 [default: s1]
+    --src2-dir SRC2-DIR                 # directory beneath root with source 2 [default: s2]
+    --train-file-list TRAIN-FL          # file that contains training set [default: train.txt]
+    --valid-file-list VAL-FL            # file that contains validation set [default: validation.txt]
+                                        #
+    --num-speakers NUM-SPEAKERS         # number of speakers [default: 2]
+    --epochs EPOCHS                     # number of epochs [default: 10]
+    --batch-size BATCH-SIZE             # batch size [default: 8]
+    --num-layers NUM-LAYERS             # number of hidden layers [default: 3]
+    --layer-size LAYER-SIZE             # size of each hidden layer [default: 896]
+    --dropout DROPOUT                   # dropout rate [default: 0.5]
+                                        #
+    --hop-size HOP-SIZE                 # hop size [default: 128]
+    --window-size WINDOW-SIZE           # window size [default: 256]
+    --window WINDOW                     # window type [default: hann]
+                                        #
+    --learning-rate LEARNING-RATE       # learning rate [default: 1.0e-3]
+    --weight-decay WEIGHT-DECAY         # weight decay [default: 1.0e-5]
+    --factor FACTOR                     # factor [default: 0.7]
+    --patience PATIENCE                 # patience [default: 0]
+    --min-learning-rate MIN-LEARNING-RATE
+                                        # min-learning-rate [default: 1.0e-10]
+    --clip-norm CLIP-NORM               # clip norm [default: 200]
+
+Examples:
+    ./train.py -d --epochs 2
+"""
+
+from docopt import docopt
+
+from itertools import permutations
+
+import librosa
+
+import torch as th
[…]

### 2018-12-15 code/upit/train.py
upit full setup.

@@ -1,0 +1,577 @@
+#!/usr/bin/env python3
+# train.py
+"""Train a uPIT BLSTM
+
+Usage:
+    train.py [options]
+
+Options:
+    -d, --debug                         # output extra debug information
+    --debug-level LEVEL                 # debug level to LEVEL [default: 1]
+                                        #
+    --train-root-dir TRAIN-ROOT-DIR     # training set root directory [default: data/2spk8kmax/tr]
+    --valid-root-dir VALID-ROOT-DIR     # validation set root directory [default: data/2spk8kmax/cv]
+    --mix-dir MIX-DIR                   # directory beneath root with mixtures [default: mix]
+    --src1-dir SRC1-DIR                 # directory beneath root with source 1 [default: s1]
+    --src2-dir SRC2-DIR                 # directory beneath root with source 2 [default: s2]
+    --train-file-list TRAIN-FL          # file that contains training set [default: train.txt]
+    --valid-file-list VAL-FL            # file that contains validation set [default: validation.txt]
+                                        #
+    --num-speakers NUM-SPEAKERS         # number of speakers [default: 2]
+    --epochs EPOCHS                     # number of epochs [default: 10]
+    --batch-size BATCH-SIZE             # batch size [default: 8]
+    --num-layers NUM-LAYERS             # number of hidden layers [default: 3]
+    --layer-size LAYER-SIZE             # size of each hidden layer [default: 896]
+    --dropout DROPOUT                   # dropout rate [default: 0.5]
+                                        #
+    --hop-size HOP-SIZE                 # hop size [default: 128]
+    --window-size WINDOW-SIZE           # window size [default: 256]
+    --window WINDOW                     # window type [default: hann]
+                                        #
+    --learning-rate LEARNING-RATE       # learning rate [default: 1.0e-3]
+    --weight-decay WEIGHT-DECAY         # weight decay [default: 1.0e-5]
+    --factor FACTOR                     # factor [default: 0.7]
+    --patience PATIENCE                 # patience [default: 0]
+    --min-learning-rate MIN-LEARNING-RATE
+                                        # min-learning-rate [default: 1.0e-10]
+    --clip-norm CLIP-NORM               # clip norm [default: 200]
+
+Examples:
+    ./train.py -d --epochs 2
+"""
+
+from docopt import docopt
+
+from itertools import permutations
+import time
+
[…]

### 2018-12-15 code/train.py
Moved into upit directory.

@@ -1,484 +1,0 @@
-#!/usr/bin/env python3
-# train.py
-"""Train a uPIT BLSTM
-
-Usage:
-    train.py [options]
-
-Options:
-    -d, --debug                         # output extra debug information
-    --debug-level LEVEL                 # debug level to LEVEL [default: 1]
-                                        #
-    --train-root-dir TRAIN-ROOT-DIR     # training set root directory [default: data/2spk8kmax/tr]
-    --valid-root-dir VALID-ROOT-DIR     # validation set root directory [default: data/2spk8kmax/cv]
-    --mix-dir MIX-DIR                   # directory beneath root with mixtures [default: mix]
-    --src1-dir SRC1-DIR                 # directory beneath root with source 1 [default: s1]
-    --src2-dir SRC2-DIR                 # directory beneath root with source 2 [default: s2]
-    --train-file-list TRAIN-FL          # file that contains training set [default: train.txt]
-    --valid-file-list VAL-FL            # file that contains validation set [default: validation.txt]
-                                        #
-    --num-speakers NUM-SPEAKERS         # number of speakers [default: 2]
-    --epochs EPOCHS                     # number of epochs [default: 10]
-    --batch-size BATCH-SIZE             # batch size [default: 8]
-    --num-layers NUM-LAYERS             # number of hidden layers [default: 3]
-    --layer-size LAYER-SIZE             # size of each hidden layer [default: 896]
-    --dropout DROPOUT                   # dropout rate [default: 0.5]
-                                        #
-    --hop-size HOP-SIZE                 # hop size [default: 128]
-    --window-size WINDOW-SIZE           # window size [default: 256]
-    --window WINDOW                     # window type [default: hann]
-                                        #
-    --learning-rate LEARNING-RATE       # learning rate [default: 1.0e-3]
-    --weight-decay WEIGHT-DECAY         # weight decay [default: 1.0e-5]
-    --factor FACTOR                     # factor [default: 0.7]
-    --patience PATIENCE                 # patience [default: 0]
-    --min-learning-rate MIN-LEARNING-RATE
-                                        # min-learning-rate [default: 1.0e-10]
-    --clip-norm CLIP-NORM               # clip norm [default: 200]
-
-Examples:
-    ./train.py -d --epochs 2
-"""
-
-from docopt import docopt
-
-from itertools import permutations
-
-import librosa
-
-import torch as th
[…]

### 2018-12-15 code/upit/train.py
Cleanup.

@@ -16,6 +16,11 @@ Options:
     --src2-dir SRC2-DIR                 # directory beneath root with source 2 [default: s2]
     --train-file-list TRAIN-FL          # file that contains training set [default: train.txt]
     --valid-file-list VAL-FL            # file that contains validation set [default: validation.txt]
+    --output-dir OUTPUT-DIR             # output directory [default: out]
+    --save-every SAVE-EVERY             # save model every N epochs [default: 10]
+    --init-model INIT-MODEL             # load initial model parameters
+    --start-epoch START-EPOCH           # count of initial epoch for continued
+                                        #   training session [default: 1]
                                         #
     --num-speakers NUM-SPEAKERS         # number of speakers [default: 2]
     --epochs EPOCHS                     # number of epochs [default: 10]
@@ -38,6 +43,15 @@ Options:
 
 Examples:
     ./train.py -d --epochs 2
+    ./train.py -d --debug-level 3 --epochs 2  --batch-size 2
+
+    # Run for 10 epochs
+    ./train.py --epochs 10
+
+    # Load previous model file and continue training
+    ./train.py --epochs 10
+    ./train.py --epochs 10 --start-epoch 11 --init-model out/model.10.pkl
+    ./train.py --epochs 10 --start-epoch 21 --init-model out/model.20.pkl
 """
 
 from docopt import docopt
@@ -45,11 +59,13 @@ from docopt import docopt
 from itertools import permutations
 import time
 
+from pathlib import Path
+
 import librosa
 
-import torch as th
 import numpy as np
 
+import torch as th
 from torch.optim.lr_scheduler import ReduceLROnPlateau
 from torch.nn.utils.rnn import pack_sequence, pad_sequence
 from torch.nn.utils.rnn import PackedSequence, pad_packed_sequence
@@ -132,6 +148,7 @@ class TrainUpit(object):
                 f"Window: {self.window}")
 
         self.epochs = int(args['--epochs'])
+        self.start_epoch = int(args['--start-epoch'])
         self.num_layers = int(args['--num-layers'])
         self.layer_size = int(args['--layer-size'])
         self.batch_size = int(args['--batch-size'])
@@ -186,38 +203,11 @@ class TrainUpit(object):
             batch_set = sorted([valid_data[i] for i in batch_indices],
                     key=lambda x: x[0].shape[0],
                     reverse=True)
-            dprint(f"Batch indices: {batch_indices}", level=2)
+            dprint(f"Batch indices: {batch_indices}", level=3)
 
[…]

### 2018-12-15 paper/main.tex

@@
 \begin{abstract}
-In this project we evaluated and implemented state of the art techniques to solve the monaural speech separation (diarization) problem using a variety of approached.
+In this project we evaluated and implemented state of the art techniques to solve the monaural speech separation (diarization) problem using a variety of approaches.
 We investigated classification, masking, and deep learning approaches to separate a generated set of mixed speech recordings.
@@
 This task is particularly challenging where the input is monophonic, there is no control over environmental conditions, the number of speakers is unknown and there are no samples of the speakers from which to train.
-Using the CSR-I(WSJ0) Complete dataset\cite{garofolo:wsj0}, we tested several algorithms for accomplishing monaural speech separation including classification (KMeans), various masking approaches (Nonnegative Matrix Factorization [NMF], Ideal Ratio Mask[IRM], and complex Ideal Ratio Mask [cIRM]), and permutation-based deep learning approaches (Permutation Invariant Training [PIT] and utterance-level Permutation Invariant Training [uPIT]).
+Using the wsj0-mix dataset \cite{wsjmix}, we tested several algorithms for accomplishing monaural speech separation including classification ($k$-means), various masking approaches (Nonnegative Matrix Factorization [NMF], Ideal Ratio Mask [IRM], and complex Ideal Ratio Mask [cIRM]), and permutation-based deep learning approaches (Permutation Invariant Training [PIT] and utterance-level Permutation Invariant Training [uPIT]).
 
@@
 
-There are several approaches to signal separation (e.g., denoising, diarization) based on the general concept that audio can be represented as the combination of a frequency matrix and a corresponding matrix temporal activations.
+There are several approaches to signal separation (e.g., denoising, diarization) based on the general concept that audio can be represented as the combination of a frequency matrix and a corresponding matrix of temporal activations.
 As a supervised learning problem, a time-frequency  representation can be used to emphasize the points at which a target signal dominates the mixed audio and deemphasize the points where an alternate speaker (or other noise source) predominates.\cite{DBLP:journals/corr/abs-1708-07524}
@@
 More recently, probabilistic approaches have been used to represent the signal as a collection of frequency spectrum (hidden states) over time.
-On such approach, Nonnegative Hidden Markov Model (N-HMM), models a magnitude spectrogram for each latent component (source) at a given time frame as one of many states; like a traditional HMM, movement between states - and the spectral representation governing each latent component - is learned and defined by a transition matrix.
+One such approach, Nonnegative Hidden Markov Model (N-HMM), models a magnitude spectrogram for each latent component (source) at a given time frame as one of many states; like a traditional HMM, movement between states - and the spectral representation governing each latent component - is learned and defined by a transition matrix.
 \cite{10.1007/978-3-642-15995-4_18}
@@
 
-A seminal paper in applying deep learning to the task of speaker diarization was the Deep Clustering (DPCL) \cite{DBLP:journals/corr/HersheyCRW15} approach proposed by Hershey, et.
-al.
-They divided the task of separating source signals into two steps.
+A seminal paper in applying deep learning to the task of speaker diarization was the Deep Clustering (DPCL) \cite{DBLP:journals/corr/HersheyCRW15} approach proposed by Hershey, et al.
+This paper divides the task of separating source signals into two steps.
 The first step is to use a deep neural network to learn a set of embeddings that estimate a class-independent, low-rank pairwise affinity matrix.
-These embeddings are trained to minimize the distance between embeddings in the same partition while maximizing the distances between embeddings in different partitions.
+These embeddings are trained to minimize the distance between embeddings in the same speaker partition while maximizing the distances between embeddings in different partitions.
 The partitions used in their technique do not include class labels so the model does not assign a particular class to each embedding.
@@
 This is in contrast to previous work that relies on \textit{spectral clustering} for segmentation which uses local affinity measures to optimize an objective function using spectral decomposition.
-The approach of Hershey, et.
-al.
-is typical in current research for deep learning: rather than using a complicated set of specially designed features, the deep clustering approach uses a deep neural network to discover the best features for producing the desired partitions.
+The approach of Hershey, et al.
+is typical of current research for deep learning: rather than using a complicated set of specially designed features, the deep clustering approach uses a deep neural network to discover the best features for producing the desired partitions.
 
 This work is also important because it introduced a new training dataset for the task of monaural speech separation.
-The authors found that existing datasets were too limited in size or were not well fitted to their task.
-The deep learning techniques also require a large training set and so the authors adapted the Wall Street Journal (WSJ0) \cite{garofolo:wsj0} single-channel speaker corpus to produce a new dataset called WSJ0-2mix \cite{WSJmix} that was used by the later literature for comparison with DPCL.
+The authors found that existing datasets were too limited in size or were not well suited to their task.
+Deep learning techniques also require a large training set and so the authors adapted the CSR-I (WSJ0) Complete \cite{garofolo:wsj0} single-channel speaker corpus to produce a new dataset called wsj0-mix \cite{wsjmix} that was used by the later literature for comparison with DPCL.
 
-The Deep Clustering technique achieved Signal-to-Distortion (SDR) improvements of up to 6 dB which outperformed their baseline of an oracle NMF approach of 5 dB SDR which was trained on the clean speaker sources.
-It is worth noting the these initial results were improved to 10.3 dB by further optimization of the model by using a deeper architecture, a longer temporal context, and improving the regularization of the network by \cite{DBLP:journals/corr/IsikRCWH16}.
+DPCL achieved Signal-to-Distortion (SDR) improvements of up to 6 dB which outperformed their baseline oracle NMF approach of 5 dB SDR which was trained on the clean speaker sources.
+It is worth noting that these initial results were improved to 10.3 dB by further optimization of the model by using a deeper architecture, a longer temporal context, and improving the regularization of the network by Isik, et al.
+called DPCL+ \cite{DBLP:journals/corr/IsikRCWH16}.
 This advancement shows the improvements possible by tuning hyper-parameters of a deep learning model when applied to the task of speaker diarization.
@@
 Note that DPCL, discussed above, avoids this problem due to its treatment of the source separation problem as a partitioning problem.
-Although DPCL was successful, it made a simplifying assumption that each time-frequency bin was assigned to a single speaker (determined by the results of clustering).
-In \cite{DBLP:journals/corr/YuKTJ16} Yu, et.
-al.
-proposed a new technique called permutation invariant training (PIT) that avoids the label ambiguity problem by treating the reference source streams as a set rather than an ordered list.
-One key innovation of PIT is to directly train on all possible permutations of sources ($S!$ possible assignments) and to compute the combined pairwise MSE for every assignment between each reference and the estimated source.
-PIT then directly minimized the MSE of the assignment with the lowest total MSE of every permutation.
-This means that the model is directly trained on label assignment and error evaluation.
+In \cite{DBLP:journals/corr/YuKTJ16} Yu, et al.
+proposed a new technique called permutation invariant training (PIT) that avoids the label permutation problem by treating the reference source streams as a set rather than an ordered list.
+PIT trains the network on all possible permutations of sources ($S!$ possible assignments) and computes the combined pairwise MSE for every assignment between each reference and the estimated source.
+PIT then directly minimizes the MSE of the assignment with the lowest total MSE of every permutation.
+This means that the model is learns the ideal label assignment for each meta-frame.
 
@@
 Optimal speaker assignment means that as a post-processing step the speaker assignment from PIT is improved by tracing the speakers across meta-frames and determining the optimal speaker assignment for each frame.
-This improves performance because the PIT method assumes that the output-speaker does not switch across frames.
+This improves performance because PIT assumes that the output-speaker does not switch across frames.
 Without the optimal assignment the performance of PIT degrades from a high of 10.9 dB to 5.6 - 7.8 dB.
 
-The authors of PIT improved their technique by proposing uPIT which is an enhancement to PIT that includes an utterance-level cost function which solves the speaker tracing problem by causing the Recurrent Neural Network (RNN) to align frames from a particular speaker to the same output stream.
+The authors of PIT further improved their technique by proposing uPIT which is an enhancement to PIT that includes an utterance-level cost function which solves the speaker tracing problem by causing the Recurrent Neural Network (RNN) to align frames from a particular speaker to the same output stream.
 uPIT (like PIT) is also a much simpler model than other proposed deep learning algorithms and the trained networks have strong generalization performance to unseen data.
@@
 This works in practice by using recurrent neural networks such as deep LSTM (long short-term memory) RNNs and bi-directional LSTM RNNs.
-Since a recurrent neural network has access to data from past frames (and in a BLSTM information both the past and future frames are available) uPIT does not need to be trained on a meta-frame of $N$ stacked frames as PIT did.
 
@@
 Using a BLSTM and a phase sensitive mask (PSM, also referred to as a cIRM - a complex ideal ratio mask) uPIT achieves results near 11 dB SDR.
-The maximum performance from the ideal PSM is about 15 dB SDR (an improvement over the 12.5 dB SDR possible with a standard IRM).
-Based on its performance and (relative) simplicity We chose uPIT as the deep learning solution that we evaluated in our project.
+The maximum performance from the ideal PSM is about 15 dB SDR (an improvement over the 12.5 dB SDR possible with a magnitude IRM).
+Based on its performance and (relative) simplicity we chose uPIT as the deep learning solution that we evaluated in our project.
 
@@
 The audio files are stored in a format called Sphere using the Shorten compression algorithm.
-The Linguistic Data Consortium website provides a small utility called sph2pipe \cite{ldc:sph2pipe} which we used to convert the files from the Sphere format to waveform format.
-We wrote a small script to automate this process and also to merge the separate disks of the WSJ0 dataset into a single folder.
+The Linguistic Data Consortium website provides a  utility called sph2pipe \cite{ldc:sph2pipe} which we used to convert the files from the Sphere format to waveform format.
 
 \subsection{WSJ0-mix}
-The WSJ0-mix dataset was developed by the authors of the DPCL \cite{DBLP:journals/corr/HersheyCRW15} paper by mixing selections from the WSJ0 training set \texttt{si\_tr\_s} and mixing them together with different levels of signal-to-noise ratios (SNR) selected from 0 dB to 5dB.
-An additional cross validation was also generated from the training data and an evaluation set was generated with the same process but using the WSJ0 evaluation set \texttt{si\_et\_05} and development set \texttt{si\_dt\_05} which use different speakers than the training set.
-We found the definition of these mixtures and MATLABS scripts \cite{WSJmix} to produce the dataset from the WSJ0 data.
-We made some small adjustments to the scripts so that they would run in Octave, a free, open-source alternative to MATLAB, and to skip the step of generating 16 kHz outputs and the 3-speaker mixtures since we would not be using them.
+The WSJ0-mix dataset was developed by the authors of the DPCL \cite{DBLP:journals/corr/HersheyCRW15} paper by mixing selections from the WSJ0 training set \texttt{si\_tr\_s} together with different levels of signal-to-noise ratios (SNR) selected from 0 dB to 5dB.
+An additional cross validation set was also generated from the training data and an evaluation set was generated with the same process using the WSJ0 evaluation set \texttt{si\_et\_05} and development set \texttt{si\_dt\_05} which use different speakers than the training set.
+We found the definition of these mixtures and MATLAB scripts \cite{wsjmix} to produce the dataset from the WSJ0 data.
 
 The scripts produce 20,000 training files, 5,000 cross validation files, and 3,000 test files.
-Each of these includes the mixed audio as well as the independent sources which have had the correct SNR adjustments made to them.
 We further selected 1,000 training examples and 100 cross-validation and testing examples since we our algorithm was not able to finish running on the full dataset.
@@
 \begin{itemize}
-    \item A chroma vector (Wikipedia) that consists of 12 feature vectors indicating how much energy of each pitch class,In the signal we used the same notation as used i Western music notation, to present every pitch class.
+    \item A chroma vector (Wikipedia) that consists of 12 feature vectors indicating how much energy of each pitch class.
+In the signal we used the same notation as used in Western music notation, to present every pitch class.
 Then use mean of each element as one of input for clustering.
-    \item MFCCs[Wikipedia] are commonly used as features in speech recognition systems, such as the systems which can automatically recognize numbers spoken into a telephone.
+    \item MFCCs (Wikipedia) are commonly used as features in speech recognition systems, such as the systems which can automatically recognize numbers spoken into a telephone.
 In this experiment, mfcc computed 13 MFCCs.The first MFCC does not represent information related to the overall shape of the spectrum.
 It only contains a constant offset.
-Therefore, here I discard the first MFCC, then I scale the MFCCs such that each coefficient dimension has zero mean and unit variance.Then use mean of MFCCs as one part of input for clustering.
+Therefore, here I discard the first MFCC, then I scale the MFCCs such that each coefficient dimension has zero mean and unit variance.
+Then use mean of MFCCs as one part of input for clustering.
     \item We divide the audio signal into smaller frames.
@@
 \subsubsection{Kmeans Clustering}
-Since we knew that the audio from call center is the conversation between 2 speakers(customer service and customer), we set k=2 for kmeans clustering.\\
+Since we knew that the audio from call center is the conversation between 2 speakers(customer service and customer), we set k=2 for kmeans clustering.
 The results for kmeans speaker diarization are shown in Figure\ref{Fig:speaker diarization 1} and Figure \ref{Fig:speaker diarization 2}.
@@
 \end{figure}
-\subsubsection{Evaluation\cite{Segmentation}}The common metric to measure how good/bad for speaker diarization experiments is the Diarization Error Rate(DER), it contains 3 main parts:
+\subsubsection{Evaluation\cite{Segmentation}}The common metric to measure how good/bad for speaker diarization experiments is the Diarization Error Rate (DER), it contains 3 main parts:
 \begin{itemize}
-    \item The first one is called Missed Speech Rate(MSR), whichi is percentage of scored time that a hypothesized non-speech segment corresponds to a reference speaker segment.
+    \item The first one is called Missed Speech Rate (MSR), which is percentage of scored time that a hypothesized non-speech segment corresponds to a reference speaker segment.
     \item The second one is False alarm speech rate (FASR), which is percentage of scored time that a hypothesized speaker is labelled as a non-speech in the reference.
-    \item The third one is Speaker error rate(SER), which is percentage of scored time that a speaker ID is assigned to the wrong speaker.
-    \item The last one is Overlap speaker rate(OSR), which is percentage of scored time that some of the multiple speakers in a segment do not get assigned to any speaker.
+    \item The third one is Speaker error rate (SER), which is percentage of scored time that a speaker ID is assigned to the wrong speaker.
+    \item The last one is Overlap speaker rate (OSR), which is percentage of scored time that some of the multiple speakers in a segment do not get assigned to any speaker.
 \end{itemize}
@@
 
-To perform NMF, we transformed isolated audio samples of two speakers (speaker 1 and speaker 2) to a frequency-time representation using short-time Fourier transform (STFT) before decomposing each signal using the following iterative update rules:
+To perform NMF, we transformed isolated audio samples of two speakers (speaker 1 and speaker 2) to a time-frequency representation using short-time Fourier transform (STFT) before decomposing each signal using the following iterative update rules:
 
@@
 
-Using the frequency matrices associated with the sample from each speaker, a new set of activations ($W$) was generated from the mixed signal composed of both individual samples.
+Using the frequency matrices associated with the sample from each speaker, a new set of activations ($H$) was generated from the mixed signal composed of both individual samples.
 A magnitude masking matrix, reflecting the magnitudes associated with the frequency representation for each speaker at each time frame, was generated using this new set of activations and the following formula:
@@
 \begin{flushleft}
-where S1 is the basis vector associated with the speaker 1 audio sample, S2 is the basis vector associated with the speaker 2 audio sample and H is the activation matrix generated from the combined speaker 1 and speaker 2 basis vectors.\cite{ClassNMF}
+where S1 is the basis vector associated with the speaker 1 audio sample, S2 is the basis vector associated with the speaker 2 audio sample and $H$ is the activation matrix generated from the combined speaker 1 and speaker 2 basis vectors.\cite{ClassNMF}
 \end{flushleft}
@@
 
-Utterance-level Permutation Invariant Training (uPIT) is an advancement on the Permutation Invariant Training model depicted in Figure \ref{fig:pit_diagram} where the reference signals are provided to the training model as a set rather than a list by calculating the error on all permutations of the output signals and then optimizing the model on the permutation with the lowest mean squared error (MSE).
+Utterance-level Permutation Invariant Training (uPIT) is an improvement on the Permutation Invariant Training model depicted in Figure \ref{fig:pit_diagram} where the reference signals are provided to the training model as a set rather than a list by calculating the error on all permutations of the output signals and then optimizing the model on the permutation with the lowest mean squared error (MSE).
 A traditional approach is to minimize the MSE between the estimated masks and the ideal masks such as in
@@
 Our implementation of uPIT uses a BLSTM with 3 hidden layers of size 896.
-We used the Adam optimizer and set a $0.5$ rate for dropout.
-We trained using Google Colab free instances which allow using a Tesla K80 GPU with 12 Gb or RAM.
+We used the Adam optimizer and a $0.5$ rate for dropout.
+We trained using Google Colab free instances which provide a Tesla K80 GPU with 12 Gb of RAM.
 We found that training with 1,000 samples for 60 epochs took about 5 hours in this setup.
@@
 Our results with uPIT were not very good, probably due to the small training size that we used.
-We were able to find two open source implementations of uPIT one of them we used as a reference only [CITE 1] and the other [CITE 2] we were able to actually implement in order to verify our results.
+We were able to find two open source implementations of uPIT one of them we used as a reference \cite{upit2} and the other \cite{upit1} we were able to use in order to verify our results.
 The author was able to achieve results comparable to those published in the uPIT paper and he agreed that we should not expect to match the paper's performance unless we are able to train on the full dataset.
@@
 \section{Conclusion}
-We got our best results from the cIRM with NMF separation. uPIT needed more training time and Kmeans...
+Through our research we learned about many different approaches to solving the problem of monaural speech separation and the difficulties inherent in this task, especially when clean speaker sources are not available at inference.
+We got our best results from the cIRM with NMF separation since this model was able to utilize clean speaker sources.
+The $k$-means approach was not fully tested on error rates or SDR improvement.
+Due to resource constraints we were never able to fully train and test our uPIT implementation, but it was based on completed work which achieved over 10 dB SDR improvement.
 
